### 2022-09-08 ms.tex
Initial Overleaf Import

%% Beginning of file 'sample631.tex'
%%
%% Modified 2022 May  
%%
%% This is a sample manuscript marked up using the
%% AASTeX v6.31 LaTeX 2e macros.
%%
%% AASTeX is now based on Alexey Vikhlinin's emulateapj.cls 
%% (Copyright 2000-2015).  See the classfile for details.

%% AASTeX requires revtex4-1.cls and other external packages such as
%% latexsym, graphicx, amssymb, longtable, and epsf.  Note that as of 
%% Oct 2020, APS now uses revtex4.2e for its journals but remember that 
%% AASTeX v6+ still uses v4.1. All of these external packages should 
%% already be present in the modern TeX distributions but not always.
%% For example, revtex4.1 seems to be missing in the linux version of
%% TexLive 2020. One should be able to get all packages from www.ctan.org.
%% In particular, revtex v4.1 can be found at 
%% https://www.ctan.org/pkg/revtex4-1.

%% The first piece of markup in an AASTeX v6.x document is the \documentclass
%% command. LaTeX will ignore any data that comes before this command. The 
%% documentclass can take an optional argument to modify the output style.
%% The command below calls the preprint style which will produce a tightly 
%% typeset, one-column, single-spaced document.  It is the default and thus
%% does not need to be explicitly stated.
%%
%% using aastex version 6.3
\documentclass[linenumbers]{aastex631}

%% The default is a single spaced, 10 point font, single spaced article.
%% There are 5 other style options available via an optional argument. They
%% can be invoked like this:
%%
%% \documentclass[arguments]{aastex631}
%% 
%% where the layout options are:
%%
%%  twocolumn   : two text columns, 10 point font, single spaced article.
%%                This is the most compact and represent the final published
%%                derived PDF copy of the accepted manuscript from the publisher
%%  manuscript  : one text column, 12 point font, double spaced article.
%%  preprint    : one text column, 12 point font, single spaced article.  
%%  preprint2   : two text columns, 12 point font, single spaced article.
%%  modern      : a stylish, single text column, 12 point font, article with
%% 		  wider left and right margins. This uses the Daniel
%% 		  Foreman-Mackey and David Hogg design.
%%  RNAAS       : Supresses an abstract. Originally for RNAAS manuscripts 
%%                but now that abstracts are required this is obsolete for
%%                AAS Journals. Authors might need it for other reasons. DO NOT
%%                use \begin{abstract} and \end{abstract} with this style.
%%
%% Note that you can submit to the AAS Journals in any of these 6 styles.
%%
%% There are other optional arguments one can invoke to allow other stylistic
%% actions. The available options are:
%%
%%   astrosymb    : Loads Astrosymb font and define \astrocommands. 
%%   tighten      : Makes baselineskip slightly smaller, only works with 
%%                  the twocolumn substyle.
%%   times        : uses times font instead of the default
%%   linenumbers  : turn on lineno package.
%%   trackchanges : required to see the revision mark up and print its output
%%   longauthor   : Do not use the more compressed footnote style (default) for 
%%                  the author/collaboration/affiliations. Instead print all
%%                  affiliation information after each name. Creates a much 
%%                  longer author list but may be desirable for short 
%%                  author papers.
%% twocolappendix : make 2 column appendix.
%%   anonymous    : Do not show the authors, affiliations and acknowledgments 
%%                  for dual anonymous review.
%%
%% these can be used in any combination, e.g.
%%
%% \documentclass[twocolumn,linenumbers,trackchanges]{aastex631}
%%
%% AASTeX v6.* now includes \hyperref support. While we have built in specific
%% defaults into the classfile you can manually override them with the
%% \hypersetup command. For example,
%%
%% \hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
%%
%% will change the color of the internal links to red, the links to the
%% bibliography to green, the file links to cyan, and the external links to
%% magenta. Additional information on \hyperref options can be found here:
%% https://www.tug.org/applications/hyperref/manual.html#x1-40003
%%
%% Note that in v6.3 "bookmarks" has been changed to "true" in hyperref
%% to improve the accessibility of the compiled pdf file.
%%
%% If you want to create your own macros, you can do so
%% using \newcommand. Your macros should appear before
%% the \begin{document} command.
%%
\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}

%% Reintroduced the \received and \accepted commands from AASTeX v5.2
%\received{March 1, 2021}
%\revised{April 1, 2021}
%\accepted{\today}

%% Command to document which AAS Journal the manuscript was submitted to.
%% Adds "Submitted to " the argument.
%\submitjournal{PSJ}

%% For manuscript that include authors in collaborations, AASTeX v6.31
%% builds on the \collaboration command to allow greater freedom to 
%% keep the traditional author+affiliation information but only show
%% subsets. The \collaboration command now must appear AFTER the group
%% of authors in the collaboration and it takes TWO arguments. The last
%% is still the collaboration identifier. The text given in this
%% argument is what will be shown in the manuscript. The first argument
%% is the number of author above the \collaboration command to show with
%% the collaboration text. If there are authors that are not part of any
%% collaboration the \nocollaboration command is used. This command takes
%% one argument which is also the number of authors above to show. A
%% dashed line is shown to indicate no collaboration. This example manuscript
%% shows how these commands work to display specific set of authors 
%% on the front page.
%%
%% For manuscript without any need to use \collaboration the 
%% \AuthorCollaborationLimit command from v6.2 can still be used to 
%% show a subset of authors.
%
%\AuthorCollaborationLimit=2
%
%% will only show Schwarz & Muench on the front page of the manuscript
%% (assuming the \collaboration and \nocollaboration commands are
%% commented out).
%%
%% Note that all of the author will be shown in the published article.
%% This feature is meant to be used prior to acceptance to make the
%% front end of a long author article more manageable. Please do not use
%% this functionality for manuscripts with less than 20 authors. Conversely,
%% please do use this when the number of authors exceeds 40.
%%
%% Use \allauthors at the manuscript end to show the full author list.
%% This command should only be used with \AuthorCollaborationLimit is used.

%% The following command can be used to set the latex table counters.  It
%% is needed in this document because it uses a mix of latex tabular and
%% AASTeX deluxetables.  In general it should not be needed.
%\setcounter{table}{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% The following section outlines numerous optional output that
%% can be displayed in the front matter or as running meta-data.
%%
%% If you wish, you may supply running head information, although
%% this information may be modified by the editorial offices.
%\shorttitle{AASTeX v6.3.1 Sample article}
%\shortauthors{Schwarz et al.}
%%
%% You can add a light gray and diagonal water-mark to the first page 
%% with this command:
%% \watermark{text}
%% where "text", e.g. DRAFT, is the text to appear.  If the text is 
%% long you can control the water-mark size with:
%% \setwatermarkfontsize{dimension}
%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\graphicspath{{./}{figures/}}
%% This is the end of the preamble.  Indicate the beginning of the
%% manuscript itself with \begin{document}.

\begin{document}

\title{Impact of Mega Satellite Constellations on the Legacy Survey of Space and Time}

\author[0000-0003-2874-6464]{Peter Yoachim}
\affiliation{Department of Astronomy, University of Washington, Seattle, WA, USA}

XXX--add Meredith, Alina, Kresimir, Zeljko
\author{more authors}
\affiliation{more affilt}

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}

abstract
\end{abstract}

%% Keywords should appear after the \end{abstract} command. 
%% The AAS Journals now uses Unified Astronomy Thesaurus concepts:
%% https://astrothesaurus.org
%% You will be asked to selected these concepts during the submission process
%% but this old "keyword" functionality is maintained in case authors want
%% to include these concepts in their preprints.
\keywords{ methods: observational,  light pollution, surveys}

\section{Introduction} \label{sec:intro}

The overview paper \citep{overview}

\section{Constellations}

\bibliography{ms}{}
\bibliographystyle{aasjournal}

\end{document}

### 2022-09-28 ms.tex
draft 1

@@
-%% Beginning of file 'sample631.tex'
-%%
-%% Modified 2022 May  
-%%
-%% This is a sample manuscript marked up using the
-%% AASTeX v6.31 LaTeX 2e macros.
-%%
-%% AASTeX is now based on Alexey Vikhlinin's emulateapj.cls 
-%% (Copyright 2000-2015).  See the classfile for details.
-
-%% AASTeX requires revtex4-1.cls and other external packages such as
-%% latexsym, graphicx, amssymb, longtable, and epsf.  Note that as of 
-%% Oct 2020, APS now uses revtex4.2e for its journals but remember that 
-%% AASTeX v6+ still uses v4.1. All of these external packages should 
-%% already be present in the modern TeX distributions but not always.
-%% For example, revtex4.1 seems to be missing in the linux version of
-%% TexLive 2020. One should be able to get all packages from www.ctan.org.
-%% In particular, revtex v4.1 can be found at 
-%% https://www.ctan.org/pkg/revtex4-1.
-
-%% The first piece of markup in an AASTeX v6.x document is the \documentclass
-%% command. LaTeX will ignore any data that comes before this command. The 
-%% documentclass can take an optional argument to modify the output style.
-%% The command below calls the preprint style which will produce a tightly 
-%% typeset, one-column, single-spaced document.  It is the default and thus
-%% does not need to be explicitly stated.
-%%
 %% using aastex version 6.3
@@
 %% 		          Foreman-Mackey and David Hogg design.
-%%  RNAAS       : Supresses an abstract. Originally for RNAAS manuscripts 
+%%  RNAAS       : Suppresses an abstract. Originally for RNAAS manuscripts 
 %%                but now that abstracts are required this is obsolete for
@@
 %%
-%% AASTeX v6.* now includes \hyperref support. While we have built in specific
-%% defaults into the classfile you can manually override them with the
-%% \hypersetup command. For example,
-%%
-%% \hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
-%%
-%% will change the color of the internal links to red, the links to the
-%% bibliography to green, the file links to cyan, and the external links to
-%% magenta. Additional information on \hyperref options can be found here:
-%% https://www.tug.org/applications/hyperref/manual.html#x1-40003
-%%
-%% Note that in v6.3 "bookmarks" has been changed to "true" in hyperref
-%% to improve the accessibility of the compiled pdf file.
 %%
@@
 
-%% Reintroduced the \received and \accepted commands from AASTeX v5.2
 %\received{March 1, 2021}
@@
 %% Adds "Submitted to " the argument.
-%\submitjournal{PSJ}
+%\submitjournal{ApJL}
 
-%% For manuscript that include authors in collaborations, AASTeX v6.31
-%% builds on the \collaboration command to allow greater freedom to 
-%% keep the traditional author+affiliation information but only show
-%% subsets. The \collaboration command now must appear AFTER the group
-%% of authors in the collaboration and it takes TWO arguments. The last
-%% is still the collaboration identifier. The text given in this
-%% argument is what will be shown in the manuscript. The first argument
-%% is the number of author above the \collaboration command to show with
-%% the collaboration text. If there are authors that are not part of any
-%% collaboration the \nocollaboration command is used. This command takes
-%% one argument which is also the number of authors above to show. A
-%% dashed line is shown to indicate no collaboration. This example manuscript
-%% shows how these commands work to display specific set of authors 
-%% on the front page.
-%%
-%% For manuscript without any need to use \collaboration the 
-%% \AuthorCollaborationLimit command from v6.2 can still be used to 
-%% show a subset of authors.
-%
-%\AuthorCollaborationLimit=2
-%
-%% will only show Schwarz & Muench on the front page of the manuscript
-%% (assuming the \collaboration and \nocollaboration commands are
-%% commented out).
 %%
@@
 
-\title{Impact of Mega Satellite Constellations on the Legacy Survey of Space and Time}
+\title{Impact of Satellite Constellations on the Legacy Survey of Space and Time}
 
+\author[0000-0002-8400-1910]{Jinghan Alina Hu}
+\affiliation{Harvey Mudd College, Claremont, CA, USA}
+\author[0000-0003-1305-7308]{Meredith L. Rawls}
+\affiliation{Department of Astronomy and DiRAC, University of Washington, Seattle, WA, USA}
 \author[0000-0003-2874-6464]{Peter Yoachim}
-\affiliation{Department of Astronomy, University of Washington, Seattle, WA, USA}
-
-XXX--add Meredith, Alina, Kresimir, Zeljko
-\author{more authors}
-\affiliation{more affilt}
+\affiliation{Department of Astronomy and DiRAC, University of Washington, Seattle, WA, USA}
+\author{Zeljko Ivezic}
+\affiliation{Department of Astronomy and DiRAC, University of Washington, Seattle, WA, USA}
+\author{Kresimir Mihic}
 
@@
 \begin{abstract}
+Best abstract ever
 
-abstract
 \end{abstract}
@@
 
-The overview paper \citep{overview}
+Vera C. Rubin Observatory's Legacy Survey of Space and Time (LSST) is a ten-year
+astronomical imaging survey that will begin in 2024 from a new telescope in Chile. Instead
+of soliciting individual requests for what the telescope should point at, the LSST will uniformly
+survey the sky every three nights in six color filters to essentially create a decade-long
+high-resolution survey of the entire southern sky, and share massive quantities of data products
+with the astronomy community \citep{overview}.
 
-\section{Constellations}
+To accomplish this, there is a complex scheduler algorithm that takes into account various
+science priorities, including slew time, survey uniformity, and image depth \citep{naghib19}. The Rubin
+Scheduler algorithm creates weightings for various science priorities and uses the weighting to
+create a path for pointing observation.
+
+However, one challenge for the LSST is that increasing numbers of bright
+Low-Earth-Orbit (LEO) satellites (e.g., Starlink) are being launched, which could leave streaks
+in astronomical pointings. Over the last three years, many astronomers have raised concerns
+about the impact of this on the LEO ecosystem and astronomical surveys \citep{lawrence22,tyson20}. LEO satellites
+are visible from Earth because they reflect sunlight, especially during twilight. As the
+illuminated LEO satellites move across the field of view of an astronomical pointing, they leave
+a streak in the image, which negatively impacts the scientific value of the pointing. Astronomers
+predict there will be at least one satellite streaks in the majority of exposures of the Rubin
+Observatory \citep{lawrence22}.An example of satellite streaks in an astronomical image is shown in Figure \ref{fig-streaks}.
+
+\begin{figure}[ht!]\label{fig-streaks}
+%\includegraphics{}  % preferably .eps
+\caption{This is a caption}
+\end{figure}
+
+There have been efforts to reduce the impact of satellite streaks in astronomical pointings.
+Satellite companies like SpaceX have worked on darkening the exterior of satellites so they will
+be less illuminated [CITE SPACEX BLOG POST]. However, even with extra darkening, the Starlink satellites are still one
+magnitude too bright for the observatory cameras, meaning these darkened satellite movements
+will still be captured by the camera. Astronomers have worked on algorithms for masking
+satellite trails in images, but covering the outer rim of the trails without losing extra pixels
+remains a challenge \citep{tyson20}.
+The rapid increase in population of LEO satellites threatens to compromise the quality
+and scientific value of the LSST images and also necessitates expensive computational power to
+mask the trails. Thus, we explore another option: incorporating known commercial satellites into
+the Rubin Scheduler so the worst of them may be avoided. In this project, we create realistic
+simulated forecasts of satellite trajectories and brightness, build a tool that uses that data to create
+new scheduler constraints, and test the impact of the new scheduler on various Rubin LSST
+observing programs.
+
+%\section{Constellations}
+
+\section{Methodology}
+We begin by creating realistic forecasts of three different commercial satellite constellations: StarlinkV1 (4408 satellites, altitude 550 km), OneWeb (6372 satellites, altitude 550km), and StarlinkV2 (29988 satellites, altitude 550km).
+All three of these constellations are highly likely to launch as designed, or close to it, as both operators presently have functional satellites in orbit and are working to build and launch more.
+These simulated satellite
+constellations are illustrated in Figure \ref{fig-simulated-constellations}.
+
+\begin{figure}[ht!]\label{fig-simulated-constellations}
+%\includegraphics{}  % preferably .eps
+\caption{This is a caption}
+\end{figure}
+
+Using these simulated satellite constellations, we add an additional weighting for satellite
+dodging to the Rubin Scheduler. Given an input satellite distribution, the scheduler predicts
+satellite maps across a determined time frame. Based on the predicted location of satellites, the
+scheduler creates a reward weighting: areas with more illuminated satellites will receive a more
+negative reward weighting, or in other words, the scheduler is incentivized to avoid them. The
+scheduler then combines the satellite reward weights with other weights to determine where to
+point the telescope to take the next set of simulated observations. Figure \ref{fig-simulated-scheduler} shows an example of
+scheduler weighting for the three simulated satellite constellations at two different sun altitudes.
+
+\begin{figure}[ht!]\label{fig-simulated-scheduler}
+%\includegraphics{}  % preferably .eps
+\caption{This is a caption}
+\end{figure}
+
+As shown in Figure \ref{fig-simulated-scheduler}, the predicted OneWeb constellation map has more negative reward
+weighting than other constellations. Although the OneWeb constellation has fewer satellites than
+the StarlinkV2 constellation, the OneWeb constellation is at a higher altitude, meaning that its
+satellites will be illuminated for a longer portion of the night, and not just close to twilight.
+Therefore, OneWeb shows more negative weighting than StarlinkV2 though it has a smaller
+satellite population.
+To investigate whether the scheduler behaves how we expect with the new satellite
+weighting added we create a testing function that measures the length of satellite streaks in the
+simulated field of views. To ensure efficiency, only satellites that are above the altitude limit and
+illuminated by the sun will be considered. Satellites below the altitude limit (indicated in the gray
+region in Figure \ref{fig-simulated-scheduler}) can’t be observed in pointings and are therefore not included for efficiency.
+For each satellite, we first determine whether they are in the field of view for any given
+telescope pointing by calculating their distance from the center of the field of view. If their
+distance from the field of view is less than the radius of the field of view, the satellite has crossed
+through the pointing. To quantify the impact of the satellite on the pointing, we then project the satellite location as well as the pointing to a 2D x,y plane. In the 2D plane, the field of pointing
+becomes roughly a circle and the start and end locations of the satellite become two points on the
+plane. Using the shapely python library, we calculate the intersection length of the satellite location
+points and the circle.
+Therefore, with a given simulated satellite constellation and a schedule of observations,
+we record the number of satellites that occur in each field of view and measure how long the
+satellite streaks are. Using a realistic estimated satellite streak width of 50 pixels, we can convert
+the streak length results into the number of affected image pixels, which allows us to
+quantitatively measure the efficiency of the dodging algorithm with pixel loss.
+
+\section{Results}
+We begin by investigating the efficiency of the added weighting algorithm on dodging
+satellites in observations. As shown in Figure \ref{fig-pixel-loss-weight}, using a conservative estimate of 50 pixels as
+streak width, we found that higher dodging weights reduces pixels lost to satellite streaks. We
+also found that smaller constellations at lower orbital altitudes (StarlinkV1 for example)
+inherently results in less pixel loss per pointing, nearly independent of the dodging weight.
+Figure \ref{fig-pixel-loss-weight} shows that the added weighted dodging algorithm was able to effectively avoid satellite streaks in pointings.
+
+\begin{figure}[ht!]\label{fig-pixel-loss-weight}
+%
+%\includegraphics{}  % preferably .eps
+\caption{This is a caption}
+\end{figure}
+
+We then investigate the relationship between number of exposures taken and the dodging
+weight. As shown in Figure \ref{fig-pixel-loss-weight}, higher dodging weight results in fewer exposures, most likely due
+to longer slew times. With the additional avoidance weighting, the telescope may be prompted to
+slew to a location other than the desired pointing location and then slew back in order to avoid
+some satellites, resulting in fewer overall exposures. We also found that larger constellations like
+StarlinkV2 tend to decrease the number of exposures slightly more than smaller constellations
+(StarlinkV1), which is expected. More satellites or satellites at higher orbital altitudes result in
+bigger areas of avoidance on the sky, which results in more slewing to avoid the affected areas,
+which subsequently reduces exposures. \\
+
+We now investigate the tradeoff between survey depth and satellite avoidance. One goal of LSST is to collect a large number of exposures of all parts of the southern sky so the
+exposures can be coadded to reveal faint structures that aren’t visible in individual images.
+Therefore, survey depth is crucial to LSST science goals and the trade off must be evaluated.
+With the satellite avoidance algorithm, the scheduler is prompted to avoid regions with
+illuminated satellites, which could result in longer slew time or less desirable pointing
+conditions, which contribute to a loss in survey depth. Therefore, we evaluate the tradeoff
+between survey depth loss and satellite avoidance using the dodging algorithm.\\
+
+\begin{figure}[ht!]\label{fig-trade-off}
+%\includegraphics{}  % preferably .eps
+\caption{This is a caption}
+\end{figure}
+
+As shown in Figure \ref{fig-trade-off} , for all three constellations, minor reduction in mean pixel loss per visit
+($0.001\%$ decrease) results in a significant decrease in median coadded depth. Currently, pixel
+loss due to satellite streaks only constitutes a very low percentage of total pixels in a pointing
+(about $0.008\%$ without any dodging). Given that the percentage is currently so small, a very
+minor reduction in mean pixel loss per visit results in significant reduction in image depth.
+
+\section{Discussion and Conclusion}
+
+We have demonstrated that adding a weighted term in the scheduler algorithm for
+illuminated satellites can effectively reduce the amount of satellite streaks in observations, and
+subsequently reduce mean pixel loss per visit. This is shown in Figure \ref{fig-pixel-loss-weight} where all three
+constellations caused less pixel loss with increasing dodging weights. However, with the new
+added priority on dodging, the telescope can be pushed to take an observation path that does not
+optimize slew time, and subsequently reduces the number of exposures and overall survey depth.
+The tradeoff essentially comes down to the relationship between mean pixel loss
+reduction and survey depth reduction. If the amount of pixels saved outweighs the amount of
+survey depth lost, then the algorithm is valuable to be added to the scheduler. The tradeoff
+relationship is evaluated in Figure \ref{fig-trade-off}. We found that, for all three constellations, the tradeoff
+between reduction of pixel loss and loss in coadded image depth is so significant that satellite
+dodging is overwhelmingly not worth it, provided the simulated constellations we selected are
+representative of reality. Since each pointing contains such a large number of pixels, the number
+of pixels destroyed by satellite streaks constitutes a very small percentage of the overall image.
+Therefore, while the weighted dodging algorithm is effective at reducing satellite streaks in
+images, at the current stage, satellite avoidance is mostly unnecessary considering the trade-off
+between pixel loss and loss in overall survey depth.
+
+\section{Future Work}
+While the satellite dodging algorithm is mostly unnecessary for the three satellite
+constellations we considered, many astronomers estimate that there will be sharp increases in
+satellite population in the next 5–10 years, which notably overlaps the LSST operations period
+(2024–2035). With the dramatic increase in satellite population, the dodging algorithm might
+become more relevant. In addition, satellites from other operators may be significantly brighter
+than present-day Starlink and OneWeb satellites, and may saturate the LSST Camera’s detectors.
+In this scenario, dodging may be required to salvage some science, even if it means losing out on
+survey depth.
+Related to this, one future work direction involves adding in brightness weighting to the
+dodging algorithm. The idea is to only avoid satellites brighter than a certain brightness
+threshold. This could potentially reduce the region of avoidance, therefore reducing the loss in
+coadded depth — it could reduce the tradeoff between pixel loss and loss in survey depth,
+making this preliminary work directly relevant for a successful LSST. Finally, it may be possible
+to compute optimal starting locations for observation schedules based on predicted satellite maps
+in order to optimize satellite avoidance.
+
+\software{Astropy \citep{astropy2013, astropy2018}, 
+          Healpy and HEALpix\footnote{\url{http://healpix.sourceforge.net}} \citep{healpix2005, Zonca2019},
+          Matplotlib \citep{Hunter:2007},
+          Numpy \citep{harris2020array},
+          rubin\_sim \citep{yoachim2022},
+          Scipy \citep{2020SciPy-NMeth},
+          Shapely\footnote{https://github.com/shapely/shapely}}
+          % todo: add citation to newest 2022 astropy paper
 

### 2022-09-28 ms.tex

@@
 predict there will be at least one satellite streaks in the majority of exposures of the Rubin
-Observatory \citep{lawrence22}.An example of satellite streaks in an astronomical image is shown in Figure \ref{fig-streaks}.
+Observatory \citep{lawrence22}.
+% XXX-but, our results say OneWeb streaks 10% of Rubin images.
+We shouldn't repeat claims we know are wrong.
+And this is just citing tyson20, which does not say the majority of exposures would have a streak.
+An example of satellite streaks in an astronomical image is shown in Figure \ref{fig-streaks}.
 
-\begin{figure}[ht!]\label{fig-streaks}
-%\includegraphics{}  % preferably .eps
-\caption{This is a caption}
+\begin{figure}[ht!]
+%\includegraphics{} 
+\caption{Example image of a satellite streak \label{fig-streaks}}
 \end{figure}
@@
 \section{Methodology}
-We begin by creating realistic forecasts of three different commercial satellite constellations: StarlinkV1 (4408 satellites, altitude 550 km), OneWeb (6372 satellites, altitude 550km), and StarlinkV2 (29988 satellites, altitude 550km).
-All three of these constellations are highly likely to launch as designed, or close to it, as both operators presently have functional satellites in orbit and are working to build and launch more.
+We begin by creating realistic forecasts of three different commercial satellite constellations: StarlinkV1 (4408 satellites, altitude 540-570 km), OneWeb (6372 satellites, altitude 1200 km), and StarlinkV2 (29988 satellites, altitude 340-614 km).
+All three of these constellations are highly likely to launch as designed (XXX-lol, no), or close to it, as both operators presently have functional satellites in orbit and are working to build and launch more.
 These simulated satellite
@@
 
-\begin{figure}[ht!]\label{fig-simulated-constellations}
-%\includegraphics{}  % preferably .eps
-\caption{This is a caption}
+\begin{figure}[ht!]
+\epsscale{0.5}
+%% XXX--ugh, so these have enough points that they take forever to render. Thoughts on how to improve welcome
+%\plotone{plots/sats_slv1.pdf}
+%\plotone{plots/sats_slv2.pdf}
+%\plotone{plots/sats_ow.pdf}
+\epsscale{1}
+\caption{Hammer projection of the altitude and azimuth positions of our simulated satellite constellations as seen from Rubin Observatory on October 1, 2023 with a sun altitude of -18 degrees.
+Red points show un-illuminated satellites, blue points show those illuminated by the sun, and black points are illuminated and above the Rubin altitude limit.
+\label{fig-simulated-constellations}}
 \end{figure}
@@
 scheduler then combines the satellite reward weights with other weights to determine where to
-point the telescope to take the next set of simulated observations. Figure \ref{fig-simulated-scheduler} shows an example of
+point the telescope to take the next set of simulated observations. Figure~\ref{fig-simulated-scheduler} shows an example of
 scheduler weighting for the three simulated satellite constellations at two different sun altitudes.
 
-\begin{figure}[ht!]\label{fig-simulated-scheduler}
+XXX--quantify the basis function maps a bit more.
+The positions of the illuminated satellites are computed in 10s intervals.
+These maps are then summed over 90 minute blocks to generate an avoidance map.
+
+\begin{figure}[ht!]
 %\includegraphics{}  % preferably .eps
-\caption{This is a caption}
+\caption{Example of satellite avoidance maps. \label{fig-simulated-scheduler}}
 \end{figure}
@@
 
-\begin{figure}[ht!]\label{fig-pixel-loss-weight}
+\begin{figure}[ht!]
 %
 %\includegraphics{}  % preferably .eps
-\caption{This is a caption}
+\caption{Pixels lost as a function of satellite dodging weight. \label{fig-pixel-loss-weight}}
 \end{figure}
@@
 
-\begin{figure}[ht!]\label{fig-trade-off}
+\begin{figure}[ht!]
 %\includegraphics{}  % preferably .eps
-\caption{This is a caption}
+\caption{Pixel loss vs coadded depth. \label{fig-trade-off}}
 \end{figure}
@@
           Scipy \citep{2020SciPy-NMeth},
-          Shapely\footnote{https://github.com/shapely/shapely}}
+          Skyfield \citep{Rhodes2019}, 
+          Shapely \citep{shapely2007}}
           % todo: add citation to newest 2022 astropy paper

### 2022-09-29 ms.tex
Updates from Overleaf

@@
 
+XXX--I'll add some text describing the baseline scheduler that we use--note that we only run for the first year.
+Maybe add a zonodo to the repo so we can cite where we got the baseline simulation from.
+
 \begin{figure}[ht!]
@@
 \begin{figure}[ht!]
-%
-%\includegraphics{}  % preferably .eps
-\caption{Pixels lost as a function of satellite dodging weight. \label{fig-pixel-loss-weight}}
+\plottwo{plots/streaklen_v_weight.pdf}{plots/streakfrac_v_weight.pdf}
+\plottwo{plots/nexp_v_weight.pdf}{plots/streak_length_v_deltam.pdf}
+\caption{Results from simulating the first year of Rubin observations in the presence of different satellite constellations and varying how strongly the scheduler attempts to avoid satellites.
+Top panels show that, as expected, the number of streaks and mean streak length decrease as more emphasis is put on satellite dodging.
+Bottom panels show the resulting trade off, as avoiding satellites forces the scheduler to make longer slews and observe less desirable parts of the sky, resulting in fewer total exposures in the first year and a shallower median coadded depth.
+\label{fig-pixel-loss-weight}}
 \end{figure}

### 2022-09-30 ms.tex

@@
 \plottwo{plots/streaklen_v_weight.pdf}{plots/streakfrac_v_weight.pdf}
-\plottwo{plots/nexp_v_weight.pdf}{plots/streak_length_v_deltam.pdf}
+\plottwo{plots/nexp_v_weight.pdf}{plots/deltam_v_weight.pdf}
 \caption{Results from simulating the first year of Rubin observations in the presence of different satellite constellations and varying how strongly the scheduler attempts to avoid satellites.
 Top panels show that, as expected, the number of streaks and mean streak length decrease as more emphasis is put on satellite dodging.
-Bottom panels show the resulting trade off, as avoiding satellites forces the scheduler to make longer slews and observe less desirable parts of the sky, resulting in fewer total exposures in the first year and a shallower median coadded depth.
+Bottom panels show the resulting trade off, as avoiding satellites forces the scheduler to make longer slews and observe less desirable parts of the sky, resulting in fewer total exposures in the first year and a shallower median coadded depth in $g$.
+xxx--could stack these and share the x-axis. Need to fix the y-label getting cut off.
 \label{fig-pixel-loss-weight}}
@@
 
+xxx--Maybe a table for estimated pixel/area losses for different things. Something like narrow streak mask, wide streak mask, chip gaps, and bright/saturated stars.
+
 \section{Future Work}

### 2022-09-30 ms.tex

@@
 \begin{figure}[ht!]
-\epsscale{0.5}
-%% XXX--ugh, so these have enough points that they take forever to render. Thoughts on how to improve welcome
-%\plotone{plots/sats_slv1.pdf}
-%\plotone{plots/sats_slv2.pdf}
-%\plotone{plots/sats_ow.pdf}
+\epsscale{0.35}
+%% XXX--matplotlib seems to like leaving a huge bounding box buffer around 
+\plotone{plots/sats_slv1.pdf}
+\plotone{plots/sats_slv2.pdf}
+\plotone{plots/sats_ow.pdf}
 \epsscale{1}
-\caption{Hammer projection of the altitude and azimuth positions of our simulated satellite constellations as seen from Rubin Observatory on October 1, 2023 with a sun altitude of -18 degrees.
+\caption{Hammer projection of the altitude and azimuth positions of our simulated satellite constellations as seen from Rubin Observatory on October 1, 2023 (sun altitude -18 degrees).
 Red points show un-illuminated satellites, blue points show those illuminated by the sun, and black points are illuminated and above the Rubin altitude limit.

### 2022-10-04 ms.tex
Updates from Overleaf

@@
 \begin{figure}[ht!]
-%\includegraphics{}  % preferably .eps
+% \plotone{plots/}
 \caption{Example of satellite avoidance maps. \label{fig-simulated-scheduler}}
@@
 we record the number of satellites that occur in each field of view and measure how long the
-satellite streaks are. Using a realistic estimated satellite streak width of 50 pixels, we can convert
+satellite streaks are. Using a realistic estimated satellite streak width of 50 pixels ($\sim$XXX arcsec), we can convert
 the streak length results into the number of affected image pixels, which allows us to
 quantitatively measure the efficiency of the dodging algorithm with pixel loss.
+
+XXX--Should we add in the results from looking at a run with twilight NEO survey observations?
+Doesn't flow well with everything else, but I think it is a result that will help some folks calm down a bit that we aren't going to miss tons of killer asteroids even with large constellations.
 

### 2022-10-04 ms.tex
Updates from Overleaf

@@
 \begin{figure}[ht!]
-% \plotone{plots/}
-\caption{Example of satellite avoidance maps. \label{fig-simulated-scheduler}}
+\epsscale{0.3}
+\plotone{plots/starlink_tles_v1_0.00_basisfunc.pdf}
+\plotone{plots/starlink_tles_v2_0.00_basisfunc.pdf}
+\plotone{plots/oneweb_tles_0.00_basisfunc.pdf}
+\epsscale{1}
+\caption{Example of satellite avoidance maps constructed for the scheduler.
+The maps have been rotated so zenith is in the center of the image.
+By varying the weight placed on these maps, the scheduler will more actively avoid regions of the sky where satellites could streak images.
+\label{fig-simulated-scheduler}}
 \end{figure}
@@
 \begin{figure}[ht!]
-%\includegraphics{}  % preferably .eps
-\caption{Pixel loss vs coadded depth. \label{fig-trade-off}}
+\epsscale{0.35}
+\plotone{plots/slv1coadd.pdf}
+\plotone{plots/slv2coadd.pdf}
+\plotone{plots/onewcoadd.pdf}
+\epsscale{1}
+\caption{The trade off between pixels lost to streaks and final median coadded depth of the survey.
+Points to the right of the dashed line show improvement while those to the left signify the regained pixels do not adequatly compensate for the lost depth.
+In no situation are we able to prevent depth loss in at least some filters.
+\label{fig-trade-off}}
 \end{figure}

### 2022-10-04 ms.tex

@@
 
+XXX--\citet{tyson20} used a very simple algorithm to try and avoid satellite streaks.
+In this paper, rather than try to avoid all satellite streaks, we incorporate satellite avoidance as a component on the scheduler's Markov Decision Process, allowing us to avoid a large fraction of satellite streaks without significantly compromising the performance of the survey.
+
 \begin{figure}[ht!]
@@
 \epsscale{0.35}
-%% XXX--matplotlib seems to like leaving a huge bounding box buffer around 
 \plotone{plots/sats_slv1.pdf}
 \plotone{plots/sats_slv2.pdf}
-\plotone{plots/sats_ow.pdf}
+\plotone{plots/sats_ow.pdf}\\
+\plotone{plots/sats_slv1_late.pdf}
+\plotone{plots/sats_slv2_late.pdf}
+\plotone{plots/sats_ow_late.pdf}
 \epsscale{1}
@@
 Red points show un-illuminated satellites, blue points show those illuminated by the sun, and black points are illuminated and above the Rubin altitude limit.
-\label{fig-simulated-constellations}}
+\label{fig-simulated-constellations} XXX--could add a second row to show how nothing is up when the sun is very low.}
 \end{figure}

### 2022-10-05 ms.tex

@@
 
-Vera C. Rubin Observatory's Legacy Survey of Space and Time (LSST) is a ten-year
-astronomical imaging survey that will begin in 2024 from a new telescope in Chile. Instead
-of soliciting individual requests for what the telescope should point at, the LSST will uniformly
-survey the sky every three nights in six color filters to essentially create a decade-long
-high-resolution survey of the entire southern sky, and share massive quantities of data products
-with the astronomy community \citep{overview}.
+Vera C.
+Rubin Observatory's Legacy Survey of Space and Time (LSST) is a ten-year astronomical imaging survey that will begin in 2024 from a new telescope in Chile.
+Instead of soliciting individual requests for what the telescope should point at, the LSST will uniformly survey the sky every three nights in six color filters to essentially create a decade-long high-resolution survey of the entire southern sky, and share massive quantities of data products with the astronomy community \citep{overview}.
 
-To accomplish this, there is a complex scheduler algorithm that takes into account various
-science priorities, including slew time, survey uniformity, and image depth \citep{naghib19}. The Rubin
-Scheduler algorithm creates weightings for various science priorities and uses the weighting to
-create a path for pointing observation.
+To accomplish this, there is a complex scheduler algorithm that takes into account various science priorities, including slew time, survey uniformity, and image depth \citep{naghib19}.
+The Rubin Scheduler algorithm creates weightings for various science priorities and uses the weighting to create a path for pointing observation.
 
-However, one challenge for the LSST is that increasing numbers of bright
-Low-Earth-Orbit (LEO) satellites (e.g., Starlink) are being launched, which could leave streaks
-in astronomical pointings. Over the last three years, many astronomers have raised concerns
-about the impact of this on the LEO ecosystem and astronomical surveys \citep{lawrence22,tyson20}. LEO satellites
-are visible from Earth because they reflect sunlight, especially during twilight. As the
-illuminated LEO satellites move across the field of view of an astronomical pointing, they leave
-a streak in the image, which negatively impacts the scientific value of the pointing. Astronomers
-predict there will be at least one satellite streaks in the majority of exposures of the Rubin
-Observatory \citep{lawrence22}.
-% XXX-but, our results say OneWeb streaks 10% of Rubin images.
-We shouldn't repeat claims we know are wrong.
-And this is just citing tyson20, which does not say the majority of exposures would have a streak.
-An example of satellite streaks in an astronomical image is shown in Figure \ref{fig-streaks}.
+However, one challenge for the LSST is that increasing numbers of bright Low-Earth-Orbit (LEO) satellites (e.g., Starlink) are being launched, which could leave streaks in astronomical pointings.
+Over the last three years, many astronomers have raised concerns
+about the impact of this on the LEO ecosystem and astronomical surveys \citep{lawrence22,tyson20}.
+LEO satellites are visible from Earth because they reflect sunlight, especially during twilight.
+As the illuminated LEO satellites move across the field of view of an astronomical pointing, they leave a streak in the image, which negatively impacts the scientific value of the pointing.
 
@@
 
-\begin{figure}[ht!]
-%\includegraphics{} 
-\caption{Example image of a satellite streak \label{fig-streaks}}
-\end{figure}
-
 There have been efforts to reduce the impact of satellite streaks in astronomical pointings.
-Satellite companies like SpaceX have worked on darkening the exterior of satellites so they will
-be less illuminated [CITE SPACEX BLOG POST]. However, even with extra darkening, the Starlink satellites are still one
-magnitude too bright for the observatory cameras, meaning these darkened satellite movements
-will still be captured by the camera. Astronomers have worked on algorithms for masking
-satellite trails in images, but covering the outer rim of the trails without losing extra pixels
-remains a challenge \citep{tyson20}.
-The rapid increase in population of LEO satellites threatens to compromise the quality
-and scientific value of the LSST images and also necessitates expensive computational power to
-mask the trails. Thus, we explore another option: incorporating known commercial satellites into
-the Rubin Scheduler so the worst of them may be avoided. In this project, we create realistic
-simulated forecasts of satellite trajectories and brightness, build a tool that uses that data to create
-new scheduler constraints, and test the impact of the new scheduler on various Rubin LSST
-observing programs.
-
-%\section{Constellations}
+Satellite companies like SpaceX have worked on darkening the exterior of satellites so they will be less illuminated\footnote{\url{https://api.starlink.com/public-files/BrightnessMitigationBestPracticesSatelliteOperators.pdf}}.
+However, even with extra darkening, the Starlink satellites are still one magnitude too bright for the observatory cameras, meaning these darkened satellite movements will still be captured by the camera.
+Astronomers have worked on algorithms for masking satellite trails in images, but covering the outer rim of the trails without losing extra pixels remains a challenge \citep{tyson20}.
+The rapid increase in population of LEO satellites threatens to compromise the quality and scientific value of the LSST images and also necessitates expensive computational power to mask the trails.
+Thus, we explore another option: incorporating known commercial satellites into the Rubin Scheduler so the worst of them may be avoided.
+In this project, we create realistic simulated forecasts of satellite trajectories and brightness, build a tool that uses that data to create new scheduler constraints, and test the impact of the new scheduler on various Rubin LSST observing programs.
 
 \section{Methodology}
-We begin by creating realistic forecasts of three different commercial satellite constellations: StarlinkV1 (4408 satellites, altitude 540-570 km), OneWeb (6372 satellites, altitude 1200 km), and StarlinkV2 (29988 satellites, altitude 340-614 km).
+We begin by creating realistic forecasts of three different commercial satellite constellations: StarlinkV1 (4,408 satellites, altitude 540-570 km), OneWeb (6,372 satellites, altitude 1200 km), and StarlinkV2 (29,988 satellites, altitude 340-614 km).
 All three of these constellations are highly likely to launch as designed (XXX-lol, no), or close to it, as both operators presently have functional satellites in orbit and are working to build and launch more.
-These simulated satellite
-constellations are illustrated in Figure \ref{fig-simulated-constellations}.
+These simulated satellite constellations are illustrated in Figure \ref{fig-simulated-constellations}.
 
+% TODO: add three more panels on the top that show each simulated constellation
+% as it appeared in the poster
+% TODO: make label fonts on each figure GINORMOUS so they are legible when shrunk,
+% and delete any extraneous labels
 \begin{figure}[ht!]
 \epsscale{0.35}
+\plotone{plots/starlinkv1.pdf}
+\plotone{plots/starlinkv2.pdf}
+\plotone{plots/oneweb.pdf}
 \plotone{plots/sats_slv1.pdf}
@@
 Red points show un-illuminated satellites, blue points show those illuminated by the sun, and black points are illuminated and above the Rubin altitude limit.
-\label{fig-simulated-constellations} XXX--could add a second row to show how nothing is up when the sun is very low.}
+The second row is the same, but for 6 hours later when the sun has reached and altitude of -50 degrees.
+Because the Starlink constellations are lower altitude, they have no illuminated satellites in the middle of the night, while the OneWeb constellation has only one illuminated satellite above the Rubin altitude limit.
+\label{fig-simulated-constellations} }
 \end{figure}
 
-Using these simulated satellite constellations, we add an additional weighting for satellite
-dodging to the Rubin Scheduler. Given an input satellite distribution, the scheduler predicts
-satellite maps across a determined time frame. Based on the predicted location of satellites, the
-scheduler creates a reward weighting: areas with more illuminated satellites will receive a more
-negative reward weighting, or in other words, the scheduler is incentivized to avoid them. The
-scheduler then combines the satellite reward weights with other weights to determine where to
-point the telescope to take the next set of simulated observations. Figure~\ref{fig-simulated-scheduler} shows an example of
-scheduler weighting for the three simulated satellite constellations at two different sun altitudes.
+Using these simulated satellite constellations, we add an additional weighting for satellite dodging to the Rubin Scheduler.
+Given an input satellite distribution, the scheduler predicts satellite maps across a determined time frame.
+Based on the predicted location of satellites, the scheduler creates a reward weighting: areas with more illuminated satellites will receive a more negative reward weighting, or in other words, the scheduler is incentivized to avoid them.
+The scheduler then combines the satellite reward weights with other weights to determine where to point the telescope to take the next set of simulated observations.
+Figure~\ref{fig-simulated-scheduler} shows an example of scheduler weighting for the three simulated satellite constellations at two different sun altitudes.
 
@@
 \begin{figure}[ht!]
-\epsscale{0.3}
+\epsscale{0.37}
 \plotone{plots/starlink_tles_v1_0.00_basisfunc.pdf}
@@
 
-As shown in Figure \ref{fig-simulated-scheduler}, the predicted OneWeb constellation map has more negative reward
-weighting than other constellations. Although the OneWeb constellation has fewer satellites than
-the StarlinkV2 constellation, the OneWeb constellation is at a higher altitude, meaning that its
-satellites will be illuminated for a longer portion of the night, and not just close to twilight.
-Therefore, OneWeb shows more negative weighting than StarlinkV2 though it has a smaller
-satellite population.
-To investigate whether the scheduler behaves how we expect with the new satellite
-weighting added we create a testing function that measures the length of satellite streaks in the
-simulated field of views. To ensure efficiency, only satellites that are above the altitude limit and
-illuminated by the sun will be considered. Satellites below the altitude limit (indicated in the gray
-region in Figure \ref{fig-simulated-scheduler}) can’t be observed in pointings and are therefore not included for efficiency.
-For each satellite, we first determine whether they are in the field of view for any given
-telescope pointing by calculating their distance from the center of the field of view. If their
-distance from the field of view is less than the radius of the field of view, the satellite has crossed
-through the pointing. To quantify the impact of the satellite on the pointing, we then project the satellite location as well as the pointing to a 2D x,y plane. In the 2D plane, the field of pointing
-becomes roughly a circle and the start and end locations of the satellite become two points on the
-plane. Using the shapely python library, we calculate the intersection length of the satellite location
-points and the circle.
-Therefore, with a given simulated satellite constellation and a schedule of observations,
-we record the number of satellites that occur in each field of view and measure how long the
-satellite streaks are. Using a realistic estimated satellite streak width of 50 pixels ($\sim$XXX arcsec), we can convert
-the streak length results into the number of affected image pixels, which allows us to
-quantitatively measure the efficiency of the dodging algorithm with pixel loss.
-
-XXX--Should we add in the results from looking at a run with twilight NEO survey observations?
-Doesn't flow well with everything else, but I think it is a result that will help some folks calm down a bit that we aren't going to miss tons of killer asteroids even with large constellations.
+As shown in Figure \ref{fig-simulated-scheduler}, the predicted OneWeb constellation map has more negative reward weighting than other constellations.
+Although the OneWeb constellation has fewer satellites than the StarlinkV2 constellation, the OneWeb constellation is at a higher altitude, meaning that its satellites will be illuminated for a longer portion of the night, and not just close to twilight.
+Therefore, OneWeb shows more negative weighting than StarlinkV2 though it has a smaller satellite population.
+To investigate whether the scheduler behaves how we expect with the new satellite weighting added we create a testing function that measures the length of satellite streaks in the simulated field of views.
+To ensure efficiency, only satellites that are above the altitude limit and illuminated by the sun will be considered.
+Satellites below the altitude limit (indicated in the gray region in Figure \ref{fig-simulated-scheduler}) can’t be observed in pointings and are therefore not included for efficiency.
+For each satellite, we first determine whether they are in the field of view for any given telescope pointing by calculating their distance from the center of the field of view.
+If their distance from the field of view is less than the radius of the field of view, the satellite has crossed through the pointing.
+To quantify the impact of the satellite on the pointing, we then project the satellite location as well as the pointing to a 2D x,y plane.
+In the 2D plane, the field of pointing becomes roughly a circle and the start and end locations of the satellite become two points on the plane.
+Using the shapely python library, we calculate the intersection length of the satellite location points and the circle.
+Therefore, with a given simulated satellite constellation and a schedule of observations, we record the number of satellites that occur in each field of view and measure how long the satellite streaks are.
+Using a realistic estimated satellite streak width of 50 pixels ($\sim$XXX arcsec), we can convert the streak length results into the number of affected image pixels, which allows us to quantitatively measure the efficiency of the dodging algorithm with pixel loss.
 
 \section{Results}
-We begin by investigating the efficiency of the added weighting algorithm on dodging
-satellites in observations. As shown in Figure \ref{fig-pixel-loss-weight}, using a conservative estimate of 50 pixels as
-streak width, we found that higher dodging weights reduces pixels lost to satellite streaks. We
-also found that smaller constellations at lower orbital altitudes (StarlinkV1 for example)
-inherently results in less pixel loss per pointing, nearly independent of the dodging weight.
+We begin by investigating the efficiency of the added weighting algorithm on dodging satellites in observations.
+As shown in Figure \ref{fig-pixel-loss-weight}, using a conservative estimate of 50 pixels as streak width, we found that higher dodging weights reduces pixels lost to satellite streaks.
+We also found that smaller constellations at lower orbital altitudes (StarlinkV1 for example) inherently results in less pixel loss per pointing, nearly independent of the dodging weight.
 Figure \ref{fig-pixel-loss-weight} shows that the added weighted dodging algorithm was able to effectively avoid satellite streaks in pointings.
@@
 
-We then investigate the relationship between number of exposures taken and the dodging
-weight. As shown in Figure \ref{fig-pixel-loss-weight}, higher dodging weight results in fewer exposures, most likely due
-to longer slew times. With the additional avoidance weighting, the telescope may be prompted to
-slew to a location other than the desired pointing location and then slew back in order to avoid
-some satellites, resulting in fewer overall exposures. We also found that larger constellations like
-StarlinkV2 tend to decrease the number of exposures slightly more than smaller constellations
-(StarlinkV1), which is expected. More satellites or satellites at higher orbital altitudes result in
-bigger areas of avoidance on the sky, which results in more slewing to avoid the affected areas,
-which subsequently reduces exposures. \\
+We then investigate the relationship between number of exposures taken and the dodging weight.
+As shown in Figure \ref{fig-pixel-loss-weight}, higher dodging weight results in fewer exposures, most likely due to longer slew times.
+With the additional avoidance weighting, the telescope may be prompted to slew to a location other than the desired pointing location and then slew back in order to avoid some satellites, resulting in fewer overall exposures.
+We also found that larger constellations like StarlinkV2 tend to decrease the number of exposures slightly more than smaller constellations (StarlinkV1), which is expected.
+More satellites or satellites at higher orbital altitudes result in bigger areas of avoidance on the sky, which results in more slewing to avoid the affected areas, which subsequently reduces exposures.
+\\
 
-We now investigate the tradeoff between survey depth and satellite avoidance. One goal of LSST is to collect a large number of exposures of all parts of the southern sky so the
-exposures can be coadded to reveal faint structures that aren’t visible in individual images.
+We now investigate the tradeoff between survey depth and satellite avoidance.
+One goal of LSST is to collect a large number of exposures of all parts of the southern sky so the exposures can be coadded to reveal faint structures that aren’t visible in individual images.
 Therefore, survey depth is crucial to LSST science goals and the trade off must be evaluated.
-With the satellite avoidance algorithm, the scheduler is prompted to avoid regions with
-illuminated satellites, which could result in longer slew time or less desirable pointing
-conditions, which contribute to a loss in survey depth. Therefore, we evaluate the tradeoff
-between survey depth loss and satellite avoidance using the dodging algorithm.\\
+With the satellite avoidance algorithm, the scheduler is prompted to avoid regions with illuminated satellites, which could result in longer slew time or less desirable pointing conditions, which contribute to a loss in survey depth.
+Therefore, we evaluate the tradeoff between survey depth loss and satellite avoidance using the dodging algorithm.\\
 
 \begin{figure}[ht!]
-\epsscale{0.35}
-\plotone{plots/slv1coadd.pdf}
-\plotone{plots/slv2coadd.pdf}
-\plotone{plots/onewcoadd.pdf}
+\epsscale{0.7}
+%\plotone{plots/slv1coadd.pdf}
+%\plotone{plots/slv2coadd.pdf}
+%\plotone{plots/onewcoadd.pdf}\\
+\plotone{plots/streak_length_v_deltam.pdf}
 \epsscale{1}
@@
 In no situation are we able to prevent depth loss in at least some filters.
-\label{fig-trade-off}}
+\label{fig-trade-off}
+XXX--the top panels have more info, but I kinda like the lower for the simplicity. Maybe just the lower and say all the filters follow a similar trend? Or a plot for g and a plot for i?
+xxx--MR deleted the top panels, and thinks we should avoid discussing filters/bands too much because people will ask whether we are simulating sat brightness as a function of wavelength...
+which we definitely aren't}
 \end{figure}
 
-As shown in Figure \ref{fig-trade-off} , for all three constellations, minor reduction in mean pixel loss per visit
-($0.001\%$ decrease) results in a significant decrease in median coadded depth. Currently, pixel
-loss due to satellite streaks only constitutes a very low percentage of total pixels in a pointing
-(about $0.008\%$ without any dodging). Given that the percentage is currently so small, a very
-minor reduction in mean pixel loss per visit results in significant reduction in image depth.
+As shown in Figure \ref{fig-trade-off} , for all three constellations, minor reduction in mean pixel loss per visit ($0.001\%$ decrease) results in a significant decrease in median coadded depth.
+Currently, pixel loss due to satellite streaks only constitutes a very low percentage of total pixels in a pointing (about $0.008\%$ without any dodging).
+Given that the percentage is currently so small, a very minor reduction in mean pixel loss per visit results in significant reduction in image depth.
 
@@
 
-We have demonstrated that adding a weighted term in the scheduler algorithm for
-illuminated satellites can effectively reduce the amount of satellite streaks in observations, and
-subsequently reduce mean pixel loss per visit. This is shown in Figure \ref{fig-pixel-loss-weight} where all three
-constellations caused less pixel loss with increasing dodging weights. However, with the new
-added priority on dodging, the telescope can be pushed to take an observation path that does not
-optimize slew time, and subsequently reduces the number of exposures and overall survey depth.
-The tradeoff essentially comes down to the relationship between mean pixel loss
-reduction and survey depth reduction. If the amount of pixels saved outweighs the amount of
-survey depth lost, then the algorithm is valuable to be added to the scheduler. The tradeoff
-relationship is evaluated in Figure \ref{fig-trade-off}. We found that, for all three constellations, the tradeoff
-between reduction of pixel loss and loss in coadded image depth is so significant that satellite
-dodging is overwhelmingly not worth it, provided the simulated constellations we selected are
-representative of reality. Since each pointing contains such a large number of pixels, the number
-of pixels destroyed by satellite streaks constitutes a very small percentage of the overall image.
-Therefore, while the weighted dodging algorithm is effective at reducing satellite streaks in
-images, at the current stage, satellite avoidance is mostly unnecessary considering the trade-off
-between pixel loss and loss in overall survey depth.
+We have demonstrated that adding a weighted term in the scheduler algorithm for illuminated satellites can effectively reduce the amount of satellite streaks in observations, and subsequently reduce mean pixel loss per visit.
+This is shown in Figure \ref{fig-pixel-loss-weight} where all three constellations caused less pixel loss with increasing dodging weights.
+However, with the new added priority on dodging, the telescope can be pushed to take an observation path that does not optimize slew time, and subsequently reduces the number of exposures and overall survey depth.
+The tradeoff essentially comes down to the relationship between mean pixel loss reduction and survey depth reduction.
+If the amount of pixels saved outweighs the amount of survey depth lost, then the algorithm is valuable to be added to the scheduler.
+The tradeoff relationship is evaluated in Figure \ref{fig-trade-off}.
+We found that, for all three constellations, the tradeoff between reduction of pixel loss and loss in coadded image depth is so significant that satellite dodging is overwhelmingly not worth it, provided the simulated constellations we selected are representative of reality.
+Since each pointing contains such a large number of pixels, the number of pixels destroyed by satellite streaks constitutes a very small percentage of the overall image.
+Therefore, while the weighted dodging algorithm is effective at reducing satellite streaks in images, at the current stage, satellite avoidance is mostly unnecessary considering the trade-off between pixel loss and loss in overall survey depth.
+
+xxx--Add short statement about impacts on near-Earth object (NEO) searches, and also a figure that Peter will make.
+XXX--Add in the results from looking at a run with twilight NEO survey observations.
+I think it is a result that will help some folks calm down a bit that we aren't going to miss tons of killer asteroids even with large constellations.
 
@@
 
+xxx--Consider saying something like only 10\% of images will have a streak from these three constellations, whereas Lawrence et al.
+2022 says ``a majority''.
+Some of this is probably based on old Starlink info with higher altitudes.
+
 \section{Future Work}
-While the satellite dodging algorithm is mostly unnecessary for the three satellite
-constellations we considered, many astronomers estimate that there will be sharp increases in
-satellite population in the next 5–10 years, which notably overlaps the LSST operations period
-(2024–2035). With the dramatic increase in satellite population, the dodging algorithm might
-become more relevant. In addition, satellites from other operators may be significantly brighter
-than present-day Starlink and OneWeb satellites, and may saturate the LSST Camera’s detectors.
-In this scenario, dodging may be required to salvage some science, even if it means losing out on
-survey depth.
-Related to this, one future work direction involves adding in brightness weighting to the
-dodging algorithm. The idea is to only avoid satellites brighter than a certain brightness
-threshold. This could potentially reduce the region of avoidance, therefore reducing the loss in
-coadded depth — it could reduce the tradeoff between pixel loss and loss in survey depth,
-making this preliminary work directly relevant for a successful LSST. Finally, it may be possible
-to compute optimal starting locations for observation schedules based on predicted satellite maps
-in order to optimize satellite avoidance.
+While the satellite dodging algorithm is mostly unnecessary for the three satellite constellations we considered, many astronomers estimate that there will be sharp increases in satellite population in the next 5–10 years, which notably overlaps the LSST operations period (2024–2035).
+With the dramatic increase in satellite population, the dodging algorithm might become more relevant.
+In addition, satellites from other operators may be significantly brighter than present-day Starlink and OneWeb satellites, and may saturate the LSST Camera’s detectors.
+In this scenario, dodging may be required to salvage some science, even if it means losing out on survey depth.
+Related to this, one future work direction involves adding in brightness weighting to the dodging algorithm.
+The idea is to only avoid satellites brighter than a certain brightness threshold.
+This could potentially reduce the region of avoidance, therefore reducing the loss in coadded depth — it could reduce the tradeoff between pixel loss and loss in survey depth, making this preliminary work directly relevant for a successful LSST.
+Finally, it may be possible to compute optimal starting locations for observation schedules based on predicted satellite maps in order to optimize satellite avoidance.
 

### 2022-10-05 ms.tex

@@
 \epsscale{0.7}
-%\plotone{plots/slv1coadd.pdf}
-%\plotone{plots/slv2coadd.pdf}
-%\plotone{plots/onewcoadd.pdf}\\
 \plotone{plots/streak_length_v_deltam.pdf}
 \epsscale{1}
-\caption{The trade off between pixels lost to streaks and final median coadded depth of the survey.
-Points to the right of the dashed line show improvement while those to the left signify the regained pixels do not adequatly compensate for the lost depth.
-In no situation are we able to prevent depth loss in at least some filters.
-\label{fig-trade-off}
-XXX--the top panels have more info, but I kinda like the lower for the simplicity. Maybe just the lower and say all the filters follow a similar trend? Or a plot for g and a plot for i?
-xxx--MR deleted the top panels, and thinks we should avoid discussing filters/bands too much because people will ask whether we are simulating sat brightness as a function of wavelength...
-which we definitely aren't}
+\caption{The trade off between pixels lost to streaks and final median coadded depth of the survey. XXX--change y-axis to fraction of pixels. \label{fig-trade-off}}
 \end{figure}

### 2022-10-05 ms.tex

@@
 \plotone{plots/starlinkv2.pdf}
-\plotone{plots/oneweb.pdf}
+\plotone{plots/oneweb.pdf}\\
 \plotone{plots/sats_slv1.pdf}
@@
 Therefore, with a given simulated satellite constellation and a schedule of observations, we record the number of satellites that occur in each field of view and measure how long the satellite streaks are.
-Using a realistic estimated satellite streak width of 50 pixels ($\sim$XXX arcsec), we can convert the streak length results into the number of affected image pixels, which allows us to quantitatively measure the efficiency of the dodging algorithm with pixel loss.
+Using a realistic estimated satellite streak width of 50 pixels ($\sim10$ arcseconds), we can convert the streak length results into the number of affected image pixels, which allows us to quantitatively measure the efficiency of the dodging algorithm with pixel loss.
 
@@
 \epsscale{1}
-\caption{The trade off between pixels lost to streaks and final median coadded depth of the survey. XXX--change y-axis to fraction of pixels. \label{fig-trade-off}}
+\caption{The trade off between pixels lost to streaks and final median coadded depth in $g$\ for the survey.
+We assume streaks require 50 pixel wide masks (with a plate scale of 0.2 arcseconds per pixel).
+Negative changes in coadded depth represent shallower surveys.
+While the percentage of lost pixels can be decreased, it forces the scheduler to make longer slews and observe less desirable parts of the sky (e.g., higher airmass, brighter background).
+\label{fig-trade-off}}
 \end{figure}
@@
 Given that the percentage is currently so small, a very minor reduction in mean pixel loss per visit results in significant reduction in image depth.
+
+\begin{figure}
+    \centering
+    \plottwo{plots/temp}{}
+    \caption{Caption}
+    \label{fig:twi_neo}
+\end{figure}
 

### 2022-10-05 ms.tex
Updates from Overleaf

@@
 
+XXX--bottom line.
+Even in the case of multiple megaconstellations ion orbit, we expect Rubin would only lose $\sim$0.02\% of science pixels to streaks.
+Given reasonable predicitons of satellite orbits we find we can reduce losses by a factor of 2 with minimal impact to the survey science goals.
+Twilight observations at high airmass with no scheduling mitigation could see pixel losses up to 0.06\%.
 \end{abstract}
@@
     \centering
-    \plottwo{plots/temp}{}
-    \caption{Caption}
+    \plottwo{plots/temp/opsim_Count_night_night_lt_366_and_visitExposureTime_lt_20_HEAL_SkyMap.pdf}{plots/twilight_curves.pdf}
+    \caption{Impact of streaks on twilight NEO observations that can not easily be shifted to dodge satellites.}
     \label{fig:twi_neo}

### 2022-10-06 ms.tex

--- 2022-10-05 ms.tex
@@
 
+XXX--bottom line.
+Even in the case of multiple mega constellations in orbit, we expect Rubin would only lose $\sim$0.02\% of science pixels to streaks.
+Given reasonable satellite orbit predictions we find we can reduce pixel losses by a factor of 2 with minimal impact to the survey science goals.
+Even so, the tiny fraction of pixels lost makes mitigation strategies unnecessary.
+Twilight observations at high airmass with no scheduling mitigation could see pixel losses up to 0.06\%.
 \end{abstract}
@@
 However, one challenge for the LSST is that increasing numbers of bright Low-Earth-Orbit (LEO) satellites (e.g., Starlink) are being launched, which could leave streaks in astronomical pointings.
-Over the last three years, many astronomers have raised concerns
-about the impact of this on the LEO ecosystem and astronomical surveys \citep{lawrence22,tyson20}.
+Over the last three years, many astronomers have raised concerns about the impact of this on the LEO ecosystem and astronomical surveys \citep{lawrence22,tyson20}.
 LEO satellites are visible from Earth because they reflect sunlight, especially during twilight.
@@
 
-XXX--I'll add some text describing the baseline scheduler that we use--note that we only run for the first year.
-Maybe add a zonodo to the repo so we can cite where we got the baseline simulation from.
+XXX--I'll add some text describing the baseline scheduler that we use--note that we only run for the first year. As a starting point we use the baseline scheduler simulation in \citet{yoachim2022}.
 
@@
 Bottom panels show the resulting trade off, as avoiding satellites forces the scheduler to make longer slews and observe less desirable parts of the sky, resulting in fewer total exposures in the first year and a shallower median coadded depth in $g$.
-xxx--could stack these and share the x-axis. Need to fix the y-label getting cut off.
+xxx--could stack these and share the x-axis. 
 \label{fig-pixel-loss-weight}}
@@
     \centering
-    \plottwo{plots/temp}{}
-    \caption{Caption}
+    \plottwo{plots/temp/opsim_Count_night_night_lt_366_and_visitExposureTime_lt_20_HEAL_SkyMap.pdf}{plots/twilight_curves.pdf}
+    \caption{Impact of streaks on twilight NEO observations that can not easily be shifted to dodge satellites.
+Left panel shows the altitude, azimuth distribution of NEO observations in a recent survey simulation.
+The right panel shows the number of streaks that would result from different satellite constellations as a function of observation altitude.
+The legend includes an estimate of the pixel loss percentage for NEO observations assuming a 50 pixel wide mask is used.
+}
     \label{fig:twi_neo}
@@
 
+We find larger constellations will only leave streaks on $\sim10$\% of Rubin images.
+This is a much lower value than previous studies \citep{lawrence22, tyson20}.
+This is partly due to the latest Starlink plans not including higher altitude orbits that were included in the \citet{tyson} study.
+% And partly due to not conflating twilight observations with the whole survey.
+
+XXX--While satellite streaks are poised to be a nuisance to Rubin observations, it is important to remember 1) They do not saturate detectors, 2) Do not leave streaks when in earth's shadow, 3)
+
 \section{Future Work}

### 2022-10-12 ms.tex
Updates from Overleaf

@@
 \affiliation{Department of Astronomy and DiRAC, University of Washington, Seattle, WA, USA}
-\author{Kresimir Mihic}
+%% TODO: Add Kreso?
 
@@
 \section{Methodology}
-We begin by creating realistic forecasts of three different commercial satellite constellations: StarlinkV1 (4,408 satellites, altitude 540-570 km), OneWeb (6,372 satellites, altitude 1200 km), and StarlinkV2 (29,988 satellites, altitude 340-614 km).
+We begin by creating realistic forecasts of three different commercial satellite constellations: StarlinkV1 (4,408 satellites, altitude 540-570 km), OneWeb (6,372 satellites, altitude 1200 km), and Starlink Gen2 (29,988 satellites, altitude 340-614 km).
 All three of these constellations are highly likely to launch as designed (XXX-lol, no), or close to it, as both operators presently have functional satellites in orbit and are working to build and launch more.
@@
 
-XXX--I'll add some text describing the baseline scheduler that we use--note that we only run for the first year. As a starting point we use the baseline scheduler simulation in \citet{yoachim2022}.
+XXX--I'll add some text describing the baseline scheduler that we use--note that we only run for the first year. As a starting point we use the baseline scheduler simulation in \citet{yoachim2022b}.
 
@@
 \epsscale{1}
-\caption{Example of satellite avoidance maps constructed for the scheduler.
+\caption{Example of satellite avoidance maps constructed for the scheduler, at sun altitude -17.1 degrees.
 The maps have been rotated so zenith is in the center of the image.
@@
 As shown in Figure \ref{fig-simulated-scheduler}, the predicted OneWeb constellation map has more negative reward weighting than other constellations.
-Although the OneWeb constellation has fewer satellites than the StarlinkV2 constellation, the OneWeb constellation is at a higher altitude, meaning that its satellites will be illuminated for a longer portion of the night, and not just close to twilight.
-Therefore, OneWeb shows more negative weighting than StarlinkV2 though it has a smaller satellite population.
+Although the OneWeb constellation has fewer satellites than the Starlink Gen2 constellation, the OneWeb constellation is at a higher altitude, meaning that its satellites will be illuminated for a longer portion of the night, and not just close to twilight.
+Therefore, OneWeb shows more negative weighting than Starlink Gen2 though it has a smaller satellite population.
 To investigate whether the scheduler behaves how we expect with the new satellite weighting added we create a testing function that measures the length of satellite streaks in the simulated field of views.
@@
 As shown in Figure \ref{fig-pixel-loss-weight}, using a conservative estimate of 50 pixels as streak width, we found that higher dodging weights reduces pixels lost to satellite streaks.
-We also found that smaller constellations at lower orbital altitudes (StarlinkV1 for example) inherently results in less pixel loss per pointing, nearly independent of the dodging weight.
+We also found that smaller constellations at lower orbital altitudes (Starlink Gen1 for example) inherently results in less pixel loss per pointing, nearly independent of the dodging weight.
 Figure \ref{fig-pixel-loss-weight} shows that the added weighted dodging algorithm was able to effectively avoid satellite streaks in pointings.
@@
 With the additional avoidance weighting, the telescope may be prompted to slew to a location other than the desired pointing location and then slew back in order to avoid some satellites, resulting in fewer overall exposures.
-We also found that larger constellations like StarlinkV2 tend to decrease the number of exposures slightly more than smaller constellations (StarlinkV1), which is expected.
+We also found that larger constellations like Starlink Gen2 tend to decrease the number of exposures slightly more than smaller constellations (Starlink Gen1), which is expected.
 More satellites or satellites at higher orbital altitudes result in bigger areas of avoidance on the sky, which results in more slewing to avoid the affected areas, which subsequently reduces exposures.

### 2022-10-14 ms.tex
Updates from Overleaf

@@
 
-XXX--bottom line.
+We look at the impact of satellite mega constellations on the planned observations of Rubin Observatory.
+We construct and simulate the motions of planned Starlink and OneWeb satellites, and test how effectively the Rubin scheduler can avoid satellite streaks when given a satellite forecast.
 Even in the case of multiple mega constellations in orbit, we expect Rubin would only lose $\sim$0.02\% of science pixels to streaks.
-Given reasonable satellite orbit predictions we find we can reduce pixel losses by a factor of 2 with minimal impact to the survey science goals.
+Given reasonable satellite orbit predictions we find we can reduce losses by a factor of 2 with minimal impact to the survey science goals.
 Even so, the tiny fraction of pixels lost makes mitigation strategies unnecessary.
-Twilight observations at high airmass with no scheduling mitigation could see pixel losses up to 0.06\%.
+Twilight observations at high airmass with no satellite dodging mitigation could see pixel losses up to 0.06\%.
 \end{abstract}
@@
 
-To accomplish this, there is a complex scheduler algorithm that takes into account various science priorities, including slew time, survey uniformity, and image depth \citep{naghib19}.
-The Rubin Scheduler algorithm creates weightings for various science priorities and uses the weighting to create a path for pointing observation.
+To accomplish this, Rubin observatory has developed a scheduling algorithm that uses a modified Markov Decision Process which can generate lists of desirable observations in real time \citep{naghib19}.
+The Rubin scheduler balances the desire to minimize slewtime, optimize signal to noise in each image, and to maintain a uniform survey footprint.
 
-However, one challenge for the LSST is that increasing numbers of bright Low-Earth-Orbit (LEO) satellites (e.g., Starlink) are being launched, which could leave streaks in astronomical pointings.
+One challenge for the LSST is that increasing numbers of bright Low-Earth-Orbit (LEO) satellites (e.g., Starlink) are being launched, which could leave streaks in astronomical pointings.
 Over the last three years, many astronomers have raised concerns about the impact of this on the LEO ecosystem and astronomical surveys \citep{lawrence22,tyson20}.
 LEO satellites are visible from Earth because they reflect sunlight, especially during twilight.
-As the illuminated LEO satellites move across the field of view of an astronomical pointing, they leave a streak in the image, which negatively impacts the scientific value of the pointing.
+As the illuminated LEO satellites move across the field of view of an astronomical pointing, they leave a streak in the image.
+While the flux from satellite streaks can be removed, the resulting pixels typically have such low signal-to-noise as to be scientifically useless.
 
-XXX--\citet{tyson20} used a very simple algorithm to try and avoid satellite streaks.
-In this paper, rather than try to avoid all satellite streaks, we incorporate satellite avoidance as a component on the scheduler's Markov Decision Process, allowing us to avoid a large fraction of satellite streaks without significantly compromising the performance of the survey.
+\citet{tyson20} used a very simple algorithm to see if Rubin could avoid satellite streaks.
+In this paper, rather than try to avoid all satellite streaks, we incorporate satellite avoidance as a component on the scheduler's Markov Decision Process, allowing us to avoid a large fraction of satellite streaks without drastically altering the overall performance of the survey.
 
@@
 \section{Methodology}
-We begin by creating realistic forecasts of three different commercial satellite constellations: StarlinkV1 (4,408 satellites, altitude 540-570 km), OneWeb (6,372 satellites, altitude 1200 km), and Starlink Gen2 (29,988 satellites, altitude 340-614 km).
-All three of these constellations are highly likely to launch as designed (XXX-lol, no), or close to it, as both operators presently have functional satellites in orbit and are working to build and launch more.
-These simulated satellite constellations are illustrated in Figure \ref{fig-simulated-constellations}.
+
+We begin by creating realistic forecasts of three different commercial satellite constellations: Starlink Gen1 (4,408 satellites, altitude 540-570 km), OneWeb (6,372 satellites, altitude 1200 km), and Starlink Gen2 (29,988 satellites, altitude 340-614 km).
+OneWeb has successfully launched several hundred satellites, while the number of Starlink satellites is in the thousands.
+These simulated satellite constellations are illustrated in Figure~\ref{fig-simulated-constellations}.
 
@@
 \epsscale{1}
-\caption{Hammer projection of the altitude and azimuth positions of our simulated satellite constellations as seen from Rubin Observatory on October 1, 2023 (sun altitude -18 degrees).
-Red points show un-illuminated satellites, blue points show those illuminated by the sun, and black points are illuminated and above the Rubin altitude limit.
-The second row is the same, but for 6 hours later when the sun has reached and altitude of -50 degrees.
-Because the Starlink constellations are lower altitude, they have no illuminated satellites in the middle of the night, while the OneWeb constellation has only one illuminated satellite above the Rubin altitude limit.
+\caption{Top row shows the 3-D distribution of our simulated satellite constellations.
+Middle row shows Hammer projections of the altitude and azimuth positions of satellites as seen from Rubin Observatory on October 1, 2023 (sun altitude -18 degrees).
+Red points show un-illuminated satellites, blue points show those illuminated by the sun, and black points are both illuminated and above the Rubin 20 degree altitude limit.
+The bottom row is the same, but for 6 hours later when the sun has reached and altitude of -50 degrees.
+Because the Starlink constellations are lower altitude orbits, they have no illuminated satellites in the middle of the night, while the OneWeb constellation has only a single illuminated satellite above the Rubin altitude limit at this particular time.
 \label{fig-simulated-constellations} }
@@
 
-Using these simulated satellite constellations, we add an additional weighting for satellite dodging to the Rubin Scheduler.
-Given an input satellite distribution, the scheduler predicts satellite maps across a determined time frame.
-Based on the predicted location of satellites, the scheduler creates a reward weighting: areas with more illuminated satellites will receive a more negative reward weighting, or in other words, the scheduler is incentivized to avoid them.
-The scheduler then combines the satellite reward weights with other weights to determine where to point the telescope to take the next set of simulated observations.
-Figure~\ref{fig-simulated-scheduler} shows an example of scheduler weighting for the three simulated satellite constellations at two different sun altitudes.
+To simulate Rubin observations, we start with the baseline observing strategy in \citet{yoachim2022b}.
+This baseline attempts to take most observations in mixed filter pairs (e.g., $r$\ followed by $i$ 33 minutes later), and completes XXX observations in the first year.
+XXX--some cites for details of the scheduler and model observatory.
 
-XXX--quantify the basis function maps a bit more.
-The positions of the illuminated satellites are computed in 10s intervals.
-These maps are then summed over 90 minute blocks to generate an avoidance map.
-
-XXX--I'll add some text describing the baseline scheduler that we use--note that we only run for the first year. As a starting point we use the baseline scheduler simulation in \citet{yoachim2022b}.
+The baseline Rubin observing strategy uses three primary basis functions which reward (1) minimizing slewtime, (2) maximizing the depth of images (e.g., by avoiding the moon and high airmass), and (3) maintaining a uniform survey footprint.
+To this, we add a 4th basis function which rewards avoiding areas of the sky which will have high concentrations of illuminated satellites.
+Figure~\ref{fig-simulated-scheduler} shows an example of this new basis function for the three simulated satellite constellations at two different sun altitudes.
+The positions of the illuminated satellites are computed in 10s intervals and marked on the sky.
+These maps are then summed over 90 minute blocks to generate the basis function maps.
+Thus our modified scheduler does not try to dodge individual satellite streaks, but rather has a parameterized method for avoiding regions of the sky where satellite streaks are more likely.
 
@@
 
-As shown in Figure \ref{fig-simulated-scheduler}, the predicted OneWeb constellation map has more negative reward weighting than other constellations.
+As shown in Figure \ref{fig-simulated-scheduler}, the simulated OneWeb constellation has more negative area than other constellations.
 Although the OneWeb constellation has fewer satellites than the Starlink Gen2 constellation, the OneWeb constellation is at a higher altitude, meaning that its satellites will be illuminated for a longer portion of the night, and not just close to twilight.
-Therefore, OneWeb shows more negative weighting than Starlink Gen2 though it has a smaller satellite population.
+Therefore, OneWeb shows more negative regions than Starlink Gen2 though it has a smaller satellite population.
+
 To investigate whether the scheduler behaves how we expect with the new satellite weighting added we create a testing function that measures the length of satellite streaks in the simulated field of views.
@@
 Using a realistic estimated satellite streak width of 50 pixels ($\sim10$ arcseconds), we can convert the streak length results into the number of affected image pixels, which allows us to quantitatively measure the efficiency of the dodging algorithm with pixel loss.
+
+We simulate observing for only the first year of the planned Rubin 10 year survey.
+The sruvey strategy does not significantly change in later years, so satellite streak statistics should be similar for the entire survey.
+We do not include any satellite launches or deorbits, and for simplicity assume the satellite orbital parameters are constant.
+This should be a fine approximation as long as actual satellite orbital parameters are available $\sim1$\ day in advance so our avoidance basis functions can be constructed.
+Obviously, if there is no information available on mega constellation orbits, avoiding satellite streaks will be much more difficult.
 
@@
 Bottom panels show the resulting trade off, as avoiding satellites forces the scheduler to make longer slews and observe less desirable parts of the sky, resulting in fewer total exposures in the first year and a shallower median coadded depth in $g$.
-xxx--could stack these and share the x-axis. 
 \label{fig-pixel-loss-weight}}
@@
 
-As shown in Figure \ref{fig-trade-off} , for all three constellations, minor reduction in mean pixel loss per visit ($0.001\%$ decrease) results in a significant decrease in median coadded depth.
+As shown in Figure \ref{fig-trade-off}, for all three constellations, minor reduction in mean pixel loss per visit ($0.001\%$ decrease) results in a significant decrease in median coadded depth (0.02-0.2 mags).
 Currently, pixel loss due to satellite streaks only constitutes a very low percentage of total pixels in a pointing (about $0.008\%$ without any dodging).
@@
     \centering
+    % oof, I can't figure out how to force the contour plots to be rasterized. 
+    \epsscale{.8}
+    \plottwo{plots/temp/opsim_Count_night_night_lt_366_and_visitExposureTime_gt_20_HEAL_SkyMap.pdf}{plots/regular_curves.pdf}
     \plottwo{plots/temp/opsim_Count_night_night_lt_366_and_visitExposureTime_lt_20_HEAL_SkyMap.pdf}{plots/twilight_curves.pdf}
-    \caption{Impact of streaks on twilight NEO observations that can not easily be shifted to dodge satellites.
-Left panel shows the altitude, azimuth distribution of NEO observations in a recent survey simulation.
-The right panel shows the number of streaks that would result from different satellite constellations as a function of observation altitude.
-The legend includes an estimate of the pixel loss percentage for NEO observations assuming a 50 pixel wide mask is used.
-}
+    \epsscale{1}
+    \caption{Impact of streaks on a simulation including twilight NEO observations that can not easily be shifted to dodge satellites.
+Top panels show standard survey observations while bottom panels show results for observations used to search for NEOs.
+Left panels show the altitude, azimuth distribution of  observations.
+The right panels show the number of streaks that would result from different satellite constellations as a function of observation altitude.
+The legends include an estimate of the pixel loss percentage for all observations assuming a 50 pixel wide streak mask.
+No attempt to dodge satellites was made in these simulations.}
     \label{fig:twi_neo}
@@
 This is a much lower value than previous studies \citep{lawrence22, tyson20}.
-This is partly due to the latest Starlink plans not including higher altitude orbits that were included in the \citet{tyson} study.
+This is partly due to the latest Starlink plans not including higher altitude orbits that were included in the \citet{tyson20} study.
 % And partly due to not conflating twilight observations with the whole survey.
 
-XXX--While satellite streaks are poised to be a nuisance to Rubin observations, it is important to remember 1) They do not saturate detectors, 2) Do not leave streaks when in earth's shadow, 3)
+While satellite streaks are poised to be a nuisance to Rubin observations, it is important to remember they are not expected to saturate CCD detectors (thus satellite streaks can be removed with relatively narrow pixel masks) and do not leave streaks when in earth's shadow (which is most of the night).
+Thus, even if there are tens of thousands of satellites, the percentage of pixels lost is incredibly small.
+For comparison, the Rubin focal plane has a fill factor of XXX\%, meaning XXX\% of potential observations are lost to chip gaps.
+We also expect diffraction spikes from bright stars to destroy XXX\% of pixels.
 
 \section{Future Work}
+
+XXX-maybe just fold Future Work in with conclusions. I'm not sure we have a ton more future work to do as far as satellite dodging goes.
+
 While the satellite dodging algorithm is mostly unnecessary for the three satellite constellations we considered, many astronomers estimate that there will be sharp increases in satellite population in the next 5–10 years, which notably overlaps the LSST operations period (2024–2035).

### 2022-10-19 ms.tex

@@
 
-\title{Impact of Satellite Constellations on the Legacy Survey of Space and Time}
+\title{Satellite Constellation Avoidance with the Rubin Observatory Legacy Survey of Space and Time}
+%\title{Impact of Satellite Constellations on the Legacy Survey of Space and Time}
 
@@
 \affiliation{Department of Astronomy and DiRAC, University of Washington, Seattle, WA, USA}
-\author{Zeljko Ivezic}
+\author[0000-0001-5250-2633]{\v{Z}eljko Ivezi\'{c}}
 \affiliation{Department of Astronomy and DiRAC, University of Washington, Seattle, WA, USA}
@@
 
-%% Mark off the abstract in the ``abstract'' environment. 
 \begin{abstract}
-Best abstract ever
-
-We look at the impact of satellite mega constellations on the planned observations of Rubin Observatory.
-We construct and simulate the motions of planned Starlink and OneWeb satellites, and test how effectively the Rubin scheduler can avoid satellite streaks when given a satellite forecast.
-Even in the case of multiple mega constellations in orbit, we expect Rubin would only lose $\sim$0.02\% of science pixels to streaks.
-Given reasonable satellite orbit predictions we find we can reduce losses by a factor of 2 with minimal impact to the survey science goals.
-Even so, the tiny fraction of pixels lost makes mitigation strategies unnecessary.
+We examine the impact of large commercial satellite constellations in low-Earth orbit on the planned observations of the Vera C.
+Rubin Observatory Legacy Survey of Space and Time (LSST).
+By constructing and simulating the motions of planned Starlink and OneWeb satellites, we test how effectively the Rubin scheduler can avoid satellite streaks.
+With both of these satellite constellations in orbit, we expect the LSST would only lose $\sim$0.02\% of science pixels to streaks.
+Given a reasonably accurate satellite orbit forecast, we find it is possible to adjust the scheduler algorithm to reduce pixel losses by a factor of two with minimal impact to the survey science goals.
+However, the small fraction of pixels lost in a scenario with all presently planned Starlink and OneWeb satellites in orbit renders this mitigation strategy unnecessary.
 Twilight observations at high airmass with no satellite dodging mitigation could see pixel losses up to 0.06\%.
 \end{abstract}
+% TODO: add something to abstract that addresses impacts beyond main streak pixel loss?
+e.g., even if non-linear crosstalk cascades pixel loss by a factor of 5(?), the pixel loss is still (small percent here).
 
@@
 %% to include these concepts in their preprints.
-\keywords{ methods: observational,  light pollution, surveys}
+\keywords{methods: observational, light pollution, surveys}
 
@@
 Vera C.
-Rubin Observatory's Legacy Survey of Space and Time (LSST) is a ten-year astronomical imaging survey that will begin in 2024 from a new telescope in Chile.
-Instead of soliciting individual requests for what the telescope should point at, the LSST will uniformly survey the sky every three nights in six color filters to essentially create a decade-long high-resolution survey of the entire southern sky, and share massive quantities of data products with the astronomy community \citep{overview}.
+Rubin Observatory's LSST is a ten-year astronomical imaging survey that will begin in 2024 from a new telescope in Chile.
+Instead of soliciting individual requests for what the telescope should observe, the LSST will uniformly survey the sky every few nights using six color filters to essentially create a decade-long high-resolution survey of the entire southern sky, and share massive quantities of data products with the astronomy community \citep{overview}.
+To accomplish this, Rubin Observatory has developed a scheduling algorithm that uses a modified Markov Decision Process which can generate lists of desirable observations in real time \citep{naghib19}.
+The Rubin scheduler balances the desire to minimize slew time, optimize signal to noise in each image, and to maintain a uniform survey footprint.
 
-To accomplish this, Rubin observatory has developed a scheduling algorithm that uses a modified Markov Decision Process which can generate lists of desirable observations in real time \citep{naghib19}.
-The Rubin scheduler balances the desire to minimize slewtime, optimize signal to noise in each image, and to maintain a uniform survey footprint.
-
-One challenge for the LSST is that increasing numbers of bright Low-Earth-Orbit (LEO) satellites (e.g., Starlink) are being launched, which could leave streaks in astronomical pointings.
+One challenge for the LSST is that increasing numbers of bright low-Earth orbit (LEO) satellites (e.g., Starlink) are being launched, which may leave streaks in astronomical pointings.
 Over the last three years, many astronomers have raised concerns about the impact of this on the LEO ecosystem and astronomical surveys \citep{lawrence22,tyson20}.
 LEO satellites are visible from Earth because they reflect sunlight, especially during twilight.
-As the illuminated LEO satellites move across the field of view of an astronomical pointing, they leave a streak in the image.
-While the flux from satellite streaks can be removed, the resulting pixels typically have such low signal-to-noise as to be scientifically useless.
+As the illuminated satellites move across the field of view of an astronomical pointing, they leave a streak in the image.
+While the flux from satellite streaks can in many cases be identified and removed, the resulting pixels typically have such low signal-to-noise as to be scientifically useless.
+%TODO: add citations to SATCON and D&QS reports and mention some of their main takeaways in this paragraph
 
-\citet{tyson20} used a very simple algorithm to see if Rubin could avoid satellite streaks.
-In this paper, rather than try to avoid all satellite streaks, we incorporate satellite avoidance as a component on the scheduler's Markov Decision Process, allowing us to avoid a large fraction of satellite streaks without drastically altering the overall performance of the survey.
+\citet{tyson20} used a very simple algorithm to see if the LSST could avoid imaging satellite streaks.
+In this paper, rather than try to avoid all satellite streaks, we incorporate satellite avoidance as a component on the Rubin scheduler's Markov Decision Process, allowing us to avoid a large fraction of satellite streaks without drastically altering the overall performance of the survey.
 
-There have been efforts to reduce the impact of satellite streaks in astronomical pointings.
-Satellite companies like SpaceX have worked on darkening the exterior of satellites so they will be less illuminated\footnote{\url{https://api.starlink.com/public-files/BrightnessMitigationBestPracticesSatelliteOperators.pdf}}.
-However, even with extra darkening, the Starlink satellites are still one magnitude too bright for the observatory cameras, meaning these darkened satellite movements will still be captured by the camera.
-Astronomers have worked on algorithms for masking satellite trails in images, but covering the outer rim of the trails without losing extra pixels remains a challenge \citep{tyson20}.
-The rapid increase in population of LEO satellites threatens to compromise the quality and scientific value of the LSST images and also necessitates expensive computational power to mask the trails.
-Thus, we explore another option: incorporating known commercial satellites into the Rubin Scheduler so the worst of them may be avoided.
-In this project, we create realistic simulated forecasts of satellite trajectories and brightness, build a tool that uses that data to create new scheduler constraints, and test the impact of the new scheduler on various Rubin LSST observing programs.
+There are other efforts underway to reduce the impact of satellite streaks in astronomical images.
+For example, satellite companies like SpaceX have worked on darkening the exterior of satellites so they will be less illuminated\footnote{\url{https://api.starlink.com/public-files/BrightnessMitigationBestPracticesSatelliteOperators.pdf}}.
+However, even with the most effective darkening mitigations to date, satellites still appear bright to the LSST Camera, and are likely to cause effects like non-linear crosstalk that are challenging to correct in the LSST Science Pipelines and may introduce systematic biases to various science investigations.
+This is discussed in more detail in \citet{tyson20}.
+Astronomers have also worked on algorithms for masking satellite trails in images, but covering the outer rim of the trails without losing extra pixels remains a challenge \citep{hasan22}.
+The rapid increase in population of LEO satellites threatens to compromise the quality and scientific value of LSST images and also requires extra human and computer resources to effectively mask trails.
+Thus, we explore another option: incorporating known commercial satellites into the Rubin scheduler so the worst of them may be avoided.
 
-\section{Methodology}
+In this paper, we create realistic simulated forecasts of satellite trajectories and brightness in Section \ref{method}.
+We then build a tool that uses that data to create new scheduler constraints, and test the impact of the new scheduler on various Rubin LSST observing programs in Section \ref{results}.
+Finally, we discuss the resulting tradeoff between pixel loss and reduced survey depth that results from dodging satellites in Section \ref{discuss}, and lay out possibilities for future work in Section \ref{future} as the satellite population grows and changes during Rubin Operations.
+
+\section{Methodology}\label{method}
 
@@
 \epsscale{1}
-\caption{Top row shows the 3-D distribution of our simulated satellite constellations.
+\caption{Each column corresponds to one of the three simulated satellite constellations.
+Top row shows the 3-D distribution of simulated satellites around Earth.
 Middle row shows Hammer projections of the altitude and azimuth positions of satellites as seen from Rubin Observatory on October 1, 2023 (sun altitude -18 degrees).
-Red points show un-illuminated satellites, blue points show those illuminated by the sun, and black points are both illuminated and above the Rubin 20 degree altitude limit.
-The bottom row is the same, but for 6 hours later when the sun has reached and altitude of -50 degrees.
+Red points represent un-illuminated satellites, blue points show those illuminated by the sun, and black points are both illuminated and above the Rubin 20 degree altitude limit.
+The bottom row is the same Hammer  projections 6 hours later when the sun has reached an altitude of -50 degrees.
 Because the Starlink constellations are lower altitude orbits, they have no illuminated satellites in the middle of the night, while the OneWeb constellation has only a single illuminated satellite above the Rubin altitude limit at this particular time.
@@
 The maps have been rotated so zenith is in the center of the image.
+As shown in the color bar, darker regions have more illuminated satellites and therefore more negative weighting.
 By varying the weight placed on these maps, the scheduler will more actively avoid regions of the sky where satellites could streak images.
@@
 We simulate observing for only the first year of the planned Rubin 10 year survey.
-The sruvey strategy does not significantly change in later years, so satellite streak statistics should be similar for the entire survey.
+The survey strategy does not significantly change in later years, so satellite streak statistics should be similar for the entire survey.
 We do not include any satellite launches or deorbits, and for simplicity assume the satellite orbital parameters are constant.
@@
 
-\section{Results}
+\section{Results}\label{results}
+
 We begin by investigating the efficiency of the added weighting algorithm on dodging satellites in observations.
@@
 \caption{Results from simulating the first year of Rubin observations in the presence of different satellite constellations and varying how strongly the scheduler attempts to avoid satellites.
-Top panels show that, as expected, the number of streaks and mean streak length decrease as more emphasis is put on satellite dodging.
+Top panels show that, as expected, the fraction of streaked visits and mean streak length per visit decrease as more emphasis is put on satellite dodging.
 Bottom panels show the resulting trade off, as avoiding satellites forces the scheduler to make longer slews and observe less desirable parts of the sky, resulting in fewer total exposures in the first year and a shallower median coadded depth in $g$.
@@
 \epsscale{1}
-\caption{The trade off between pixels lost to streaks and final median coadded depth in $g$\ for the survey.
+\caption{The trade off between the percentage of pixels lost to streaks and final median coadded depth in $g$\ for the survey.
 We assume streaks require 50 pixel wide masks (with a plate scale of 0.2 arcseconds per pixel).
-Negative changes in coadded depth represent shallower surveys.
-While the percentage of lost pixels can be decreased, it forces the scheduler to make longer slews and observe less desirable parts of the sky (e.g., higher airmass, brighter background).
+Negative changes in coadded depth results in shallower surveys.
+While the percentage of lost pixels can be decreased, it reduces coadded depth in $g$ as it forces the scheduler to make longer slews and observe less desirable parts of the sky (e.g., higher airmass, brighter background).
 \label{fig-trade-off}}
 \end{figure}
+% TODO: more clearly scream that the takeaway here is that we can reduce pixel loss by half with minimal impact on survey depth, but pixel loss is small to begin with, so it's not worth it.
 
@@
 
-\section{Discussion and Conclusion}
+\section{Discussion}\label{discuss}
 
@@
 
-\section{Future Work}
+\section{Future Work}\label{future}
 

### 2022-10-31 ms.tex
Updates from Overleaf

--- 2022-10-12 ms.tex
@@
-%% using aastex version 6.3
-\documentclass[linenumbers]{aastex631}
+\documentclass[twocolumn]{aastex631}
 
-%% The default is a single spaced, 10 point font, single spaced article.
-%% There are 5 other style options available via an optional argument. They
-%% can be invoked like this:
-%%
 %% \documentclass[arguments]{aastex631}
@@
 %% There are other optional arguments one can invoke to allow other stylistic
-%% actions. The available options are:
+%% actions:
 %%
@@
 %%                  for dual anonymous review.
-%%
-%% these can be used in any combination, e.g.
-%%
-%% \documentclass[twocolumn,linenumbers,trackchanges]{aastex631}
-%%
-%%
-%% If you want to create your own macros, you can do so
-%% using \newcommand. Your macros should appear before
-%% the \begin{document} command.
-%%
+
 \newcommand{\vdag}{(v)^\dagger}
@@
 
-%% Command to document which AAS Journal the manuscript was submitted to.
-%% Adds "Submitted to " the argument.
 %\submitjournal{ApJL}
-
-%%
-%% Note that all of the author will be shown in the published article.
-%% This feature is meant to be used prior to acceptance to make the
-%% front end of a long author article more manageable. Please do not use
-%% this functionality for manuscripts with less than 20 authors. Conversely,
-%% please do use this when the number of authors exceeds 40.
-%%
-%% Use \allauthors at the manuscript end to show the full author list.
-%% This command should only be used with \AuthorCollaborationLimit is used.
-
-%% The following command can be used to set the latex table counters.  It
-%% is needed in this document because it uses a mix of latex tabular and
-%% AASTeX deluxetables.  In general it should not be needed.
-%\setcounter{table}{1}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%% The following section outlines numerous optional output that
-%% can be displayed in the front matter or as running meta-data.
-%%
-%% If you wish, you may supply running head information, although
-%% this information may be modified by the editorial offices.
-%\shorttitle{AASTeX v6.3.1 Sample article}
-%\shortauthors{Schwarz et al.}
-%%
-%% You can add a light gray and diagonal water-mark to the first page 
-%% with this command:
-%% \watermark{text}
-%% where "text", e.g. DRAFT, is the text to appear.  If the text is 
-%% long you can control the water-mark size with:
-%% \setwatermarkfontsize{dimension}
-%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
-%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\graphicspath{{./}{figures/}}
-%% This is the end of the preamble.  Indicate the beginning of the
-%% manuscript itself with \begin{document}.
 
@@
 
-\title{Impact of Satellite Constellations on the Legacy Survey of Space and Time}
+\title{Satellite Constellation Avoidance with the Rubin Observatory Legacy Survey of Space and Time}
 
@@
 \author[0000-0003-1305-7308]{Meredith L. Rawls}
-\affiliation{Department of Astronomy and DiRAC, University of Washington, Seattle, WA, USA}
+\affiliation{Department of Astronomy / DiRAC / Vera C. Rubin Observatory, University of Washington, Seattle, WA, USA}
 \author[0000-0003-2874-6464]{Peter Yoachim}
-\affiliation{Department of Astronomy and DiRAC, University of Washington, Seattle, WA, USA}
-\author{Zeljko Ivezic}
-\affiliation{Department of Astronomy and DiRAC, University of Washington, Seattle, WA, USA}
-%% TODO: Add Kreso?
+\affiliation{Department of Astronomy / DiRAC / Vera C. Rubin Observatory, University of Washington, Seattle, WA, USA}
+\author[0000-0001-5250-2633]{\v{Z}eljko Ivezi\'{c}}
+\affiliation{Department of Astronomy / DiRAC / Vera C. Rubin Observatory, University of Washington, Seattle, WA, USA}
 
-%% Mark off the abstract in the ``abstract'' environment. 
 \begin{abstract}
-Best abstract ever
+We investigate a novel satellite avoidance strategy to mitigate the impact of large commercial satellite constellations in low-Earth orbit on the Vera C.
+Rubin Observatory Legacy Survey of Space and Time (LSST).
+To do this, we construct and simulate the motions of planned Starlink and OneWeb satellites, test how effectively the Rubin scheduler algorithm can avoid them, and assess how the overall survey is affected.
+Given a reasonably accurate satellite orbit forecast, we find it is possible to adjust the scheduler algorithm to effectively avoid some satellites.
+Doing so would reduce pixel loss by a factor of two with minimal impact to the survey depth.
+However, the small fraction of pixels lost to streaks from Starlink and OneWeb may render this mitigation strategy unnecessary.
+With both of these satellite constellations fully deployed, and with no satellite avoidance strategy, the LSST would lose $\sim$0.02\% of science pixels to satellite streaks, assuming a typical streak width of 50 pixels, while high-airmass twilight observations would lose $\sim$0.06\% of science pixels.
+However, it is important to note that false alerts from glints and low surface brightness residuals can have more science impact than pixel loss.
+An increase in the number or brightness of satellites during Rubin Operations, together with science impacts not well-represented by pixel loss, may make implementing this satellite avoidance strategy worthwhile.
+\end{abstract}
 
-XXX--bottom line.
-Even in the case of multiple mega constellations in orbit, we expect Rubin would only lose $\sim$0.02\% of science pixels to streaks.
-Given reasonable satellite orbit predictions we find we can reduce pixel losses by a factor of 2 with minimal impact to the survey science goals.
-Even so, the tiny fraction of pixels lost makes mitigation strategies unnecessary.
-Twilight observations at high airmass with no scheduling mitigation could see pixel losses up to 0.06\%.
-\end{abstract}
+% TODO: change pixel loss takeaway in abstract to the 10% / factor of 2 thing
 
@@
 %% to include these concepts in their preprints.
-\keywords{ methods: observational,  light pollution, surveys}
+\keywords{Ground-based astronomy, Light pollution, Sky surveys, Artificial Satellites}
 
@@
 
-Vera C.
-Rubin Observatory's Legacy Survey of Space and Time (LSST) is a ten-year astronomical imaging survey that will begin in 2024 from a new telescope in Chile.
-Instead of soliciting individual requests for what the telescope should point at, the LSST will uniformly survey the sky every three nights in six color filters to essentially create a decade-long high-resolution survey of the entire southern sky, and share massive quantities of data products with the astronomy community \citep{overview}.
+Rubin Observatory's LSST is a ten-year astronomical imaging survey that will begin in 2024 from a new telescope under construction in Chile.
+Instead of soliciting individual requests for what the telescope should observe, the LSST will uniformly survey the sky every few nights using six color filters to essentially create a decade-long high-resolution survey of the entire southern sky, and share massive quantities of data products with the astronomy community \citep{overview}.
+To accomplish this, the LSST will employ a scheduling algorithm that uses a modified Markov Decision Process which can generate lists of desirable observations in real time \citep{naghib19}.
+The Rubin scheduler balances the desire to minimize slew time, optimize signal to noise in each image, and to maintain a uniform survey footprint.
 
-To accomplish this, there is a complex scheduler algorithm that takes into account various science priorities, including slew time, survey uniformity, and image depth \citep{naghib19}.
-The Rubin Scheduler algorithm creates weightings for various science priorities and uses the weighting to create a path for pointing observation.
+One challenge for the LSST is that increasing numbers of bright low-Earth orbit (LEO) satellites (e.g., Starlink) are being launched, which may leave streaks in astronomical pointings.
+LEO satellites are visible from Earth because they reflect sunlight, especially during twilight.
+As the illuminated satellites move across the field of view of an astronomical pointing, they leave a streak in the image.
+While the flux from satellite streaks can in many cases be identified and removed, the resulting pixels have much lower signal-to-noise.
+For a thorough discussion of the scientific utility of residual light after masking satellite trails, see \citet{hasan22}.
+Over the last three years, many astronomers have raised concerns about the impact of the proliferation of commercial satellites on the LEO ecosystem and astronomical surveys \citep{lawrence22,tyson20}.
+In addition, astronomers have come together with satellite operators and other stakeholders to create recommendations and strategies to mitigate impacts to observational astronomy and beyond \citep{satcon1,satcon2,dqs1,dqs2}.
 
-However, one challenge for the LSST is that increasing numbers of bright Low-Earth-Orbit (LEO) satellites (e.g., Starlink) are being launched, which could leave streaks in astronomical pointings.
-Over the last three years, many astronomers have raised concerns about the impact of this on the LEO ecosystem and astronomical surveys \citep{lawrence22,tyson20}.
-LEO satellites are visible from Earth because they reflect sunlight, especially during twilight.
-As the illuminated LEO satellites move across the field of view of an astronomical pointing, they leave a streak in the image, which negatively impacts the scientific value of the pointing.
+\citet{tyson20} used a very simple algorithm to see if the LSST could avoid imaging satellite streaks.
+They concluded that attempting to naively ``dodge'' of order 48,000 LEO satellites is useless because it is operationally inefficient.
+In this paper, rather than try to avoid all satellite streaks, we incorporate satellite avoidance as a component of the Rubin scheduler's Markov Decision Process.
+This allows us to avoid a significant fraction of satellite streaks and investigate what level of avoidance might be acceptable because it does not drastically alter the overall performance of the LSST.
 
-XXX--\citet{tyson20} used a very simple algorithm to try and avoid satellite streaks.
-In this paper, rather than try to avoid all satellite streaks, we incorporate satellite avoidance as a component on the scheduler's Markov Decision Process, allowing us to avoid a large fraction of satellite streaks without significantly compromising the performance of the survey.
+There are other efforts underway to mitigate the impact of satellite streaks in astronomical images.
+For example, satellite companies like SpaceX have worked on darkening the exterior of satellites so they will be less visible\footnote{\url{https://api.starlink.com/public-files/BrightnessMitigationBestPracticesSatelliteOperators.pdf}}.
+However, even with the most effective darkening mitigations to date, satellites still appear bright to the LSST Camera, and are likely to cause effects like non-linear crosstalk or glints that are challenging to correct with the LSST Science Pipelines software and may introduce systematic biases or spurious detections.
+This is discussed in more detail in \citet{tyson20} and on the Rubin Observatory LSST Project website\footnote{\url{https://ls.st/satcon}}.
+Astronomers have also developed algorithms for masking satellite trails in images, but covering the outer rim of the trails without losing extra pixels remains a challenge \citep{hasan22}.
+The rapid increase in population of LEO satellites threatens to compromise the quality and scientific value of LSST images and also requires extra human and computer resources to effectively mask trails.
+Thus, we explore an additional option: incorporating the orbits of known commercial satellites into the Rubin scheduler so the worst of them may be avoided.
 
-There have been efforts to reduce the impact of satellite streaks in astronomical pointings.
-Satellite companies like SpaceX have worked on darkening the exterior of satellites so they will be less illuminated\footnote{\url{https://api.starlink.com/public-files/BrightnessMitigationBestPracticesSatelliteOperators.pdf}}.
-However, even with extra darkening, the Starlink satellites are still one magnitude too bright for the observatory cameras, meaning these darkened satellite movements will still be captured by the camera.
-Astronomers have worked on algorithms for masking satellite trails in images, but covering the outer rim of the trails without losing extra pixels remains a challenge \citep{tyson20}.
-The rapid increase in population of LEO satellites threatens to compromise the quality and scientific value of the LSST images and also necessitates expensive computational power to mask the trails.
-Thus, we explore another option: incorporating known commercial satellites into the Rubin Scheduler so the worst of them may be avoided.
-In this project, we create realistic simulated forecasts of satellite trajectories and brightness, build a tool that uses that data to create new scheduler constraints, and test the impact of the new scheduler on various Rubin LSST observing programs.
+In this paper, we first create realistic simulated forecasts of satellite orbits in Section \ref{method}.
+We then build a tool that uses that data to create new scheduler constraints, and test the impact of the modified scheduler algorithm on LSST observing programs in Section \ref{results}.
+Finally, we discuss the resulting trade-off between pixel loss and reduced survey depth that results from avoiding satellites in Section \ref{discuss}, and lay out possibilities for the future as the satellite population changes during Rubin Operations.
+We also make available a GitHub repository with the data and software necessary to reproduce the paper's figures\footnote{\url{https://github.com/lsst-sims/satellite-dodging-ApJL}}.
 
-\section{Methodology}
-We begin by creating realistic forecasts of three different commercial satellite constellations: StarlinkV1 (4,408 satellites, altitude 540-570 km), OneWeb (6,372 satellites, altitude 1200 km), and Starlink Gen2 (29,988 satellites, altitude 340-614 km).
-All three of these constellations are highly likely to launch as designed (XXX-lol, no), or close to it, as both operators presently have functional satellites in orbit and are working to build and launch more.
-These simulated satellite constellations are illustrated in Figure \ref{fig-simulated-constellations}.
+\section{Methods}\label{method}
 
-% TODO: add three more panels on the top that show each simulated constellation
-% as it appeared in the poster
-% TODO: make label fonts on each figure GINORMOUS so they are legible when shrunk,
-% and delete any extraneous labels
-\begin{figure}[ht!]
+We begin by creating realistic forecasts of three commercial satellite constellations, which are illustrated in Figure~\ref{fig-simulated-constellations}.
+These are Starlink Gen1 (4,408 satellites, altitude $540-570$ km), OneWeb (6,372 satellites, altitude 1200 km), and Starlink Gen2 (29,988 satellites, altitude $340-614$ km).
+Each constellation uses orbital inclinations and number of satellite planes matching current plans\footnote{\url{https://ls.st/x1o}}.
+To date, OneWeb has launched and deployed several hundred satellites, while the number of Starlink satellites is in the thousands.
+
+\begin{figure*}[ht!]
 \epsscale{0.35}
@@
 \epsscale{1}
-\caption{Hammer projection of the altitude and azimuth positions of our simulated satellite constellations as seen from Rubin Observatory on October 1, 2023 (sun altitude -18 degrees).
-Red points show un-illuminated satellites, blue points show those illuminated by the sun, and black points are illuminated and above the Rubin altitude limit.
-The second row is the same, but for 6 hours later when the sun has reached and altitude of -50 degrees.
-Because the Starlink constellations are lower altitude, they have no illuminated satellites in the middle of the night, while the OneWeb constellation has only one illuminated satellite above the Rubin altitude limit.
-\label{fig-simulated-constellations} }
-\end{figure}
+\caption{Three simulated satellite constellations, one per column.
+The top row shows the 3D distribution of each constellation around Earth.
+The middle row shows Hammer projections of the altitude and azimuth positions of each constellation as seen from Rubin Observatory on October 1, 2023 during twilight (Sun altitude $-18$ degrees).
+Blue points are satellites illuminated by the Sun at this time, red points are satellites not illuminated by the Sun, and black points are satellites that are both illuminated and above the Rubin 20 degree altitude pointing limit.
+The bottom row is the same Hammer projections 6 hours later in the middle of the night (Sun altitude is $-50$ degrees).
+Because the Starlink satellites orbit at 550 km, none are illuminated in the middle of the night at this time of year.
+The OneWeb constellation at 1200 km has only a single illuminated satellite above the Rubin altitude limit at this particular time.
+\label{fig-simulated-constellations}
+}
+\end{figure*}
+% TODO: add number of sats in figure caption, make it clear the bottom 2 rows are instantaneous/snapshot (not integration)
 
-Using these simulated satellite constellations, we add an additional weighting for satellite dodging to the Rubin Scheduler.
-Given an input satellite distribution, the scheduler predicts satellite maps across a determined time frame.
-Based on the predicted location of satellites, the scheduler creates a reward weighting: areas with more illuminated satellites will receive a more negative reward weighting, or in other words, the scheduler is incentivized to avoid them.
-The scheduler then combines the satellite reward weights with other weights to determine where to point the telescope to take the next set of simulated observations.
-Figure~\ref{fig-simulated-scheduler} shows an example of scheduler weighting for the three simulated satellite constellations at two different sun altitudes.
+To simulate Rubin observations, we start with the baseline observing strategy in \citet{yoachim2022b}.
+This baseline attempts to take most observations in mixed filter pairs (e.g., $r$\ followed by $i$ 33 minutes later), and completes 215,000 visits in the first year.
+Rubin uses visits of one 30s exposure in $u$, and two back-to-back 15s exposures in the $grizy$ filters.
 
-XXX--quantify the basis function maps a bit more.
-The positions of the illuminated satellites are computed in 10s intervals.
-These maps are then summed over 90 minute blocks to generate an avoidance map.
+The baseline Rubin observing strategy uses three primary basis functions which reward (1) minimizing slewtime, (2) maximizing the depth of images (e.g., by avoiding the Moon and high airmass), and (3) maintaining a uniform depth survey footprint.
+To this, we add a fourth basis function which rewards avoiding areas of the sky which will have high concentrations of illuminated satellites.
+Figure~\ref{fig-simulated-scheduler} shows an example of this new basis function for the three simulated satellite constellations at two different Sun altitudes.
+The positions of the illuminated satellites are computed in 10-second intervals and marked on the sky.
+These maps are then summed over 90-minute blocks to generate the basis function maps.
+Thus our modified scheduler with a satellite avoidance strategy does not try to avoid individual satellite streaks, but rather has a parameterized method for avoiding regions of the sky where satellite streaks are more likely.
+This has an additional benefit of not requiring high precision satellite orbit forecasts.
 
-XXX--I'll add some text describing the baseline scheduler that we use--note that we only run for the first year. As a starting point we use the baseline scheduler simulation in \citet{yoachim2022b}.
-
-\begin{figure}[ht!]
+\begin{figure*}[ht!]
 \epsscale{0.37}
@@
 \epsscale{1}
-\caption{Example of satellite avoidance maps constructed for the scheduler, at sun altitude -17.1 degrees.
-The maps have been rotated so zenith is in the center of the image.
+\caption{Satellite avoidance maps constructed for the Rubin scheduler for each simulated constellation.
+Each is for a twilight observation period of 90 minutes (beginning after sunset with a Sun altitude of $-17.1$ degrees).
+The maps are rotated so zenith is in the center of the image.
+Darker regions have more illuminated satellites and therefore more negative weighting.
 By varying the weight placed on these maps, the scheduler will more actively avoid regions of the sky where satellites could streak images.
-\label{fig-simulated-scheduler}}
-\end{figure}
+\label{fig-simulated-scheduler}
+}
+\end{figure*}
 
-As shown in Figure \ref{fig-simulated-scheduler}, the predicted OneWeb constellation map has more negative reward weighting than other constellations.
-Although the OneWeb constellation has fewer satellites than the Starlink Gen2 constellation, the OneWeb constellation is at a higher altitude, meaning that its satellites will be illuminated for a longer portion of the night, and not just close to twilight.
-Therefore, OneWeb shows more negative weighting than Starlink Gen2 though it has a smaller satellite population.
-To investigate whether the scheduler behaves how we expect with the new satellite weighting added we create a testing function that measures the length of satellite streaks in the simulated field of views.
-To ensure efficiency, only satellites that are above the altitude limit and illuminated by the sun will be considered.
-Satellites below the altitude limit (indicated in the gray region in Figure \ref{fig-simulated-scheduler}) can’t be observed in pointings and are therefore not included for efficiency.
-For each satellite, we first determine whether they are in the field of view for any given telescope pointing by calculating their distance from the center of the field of view.
-If their distance from the field of view is less than the radius of the field of view, the satellite has crossed through the pointing.
-To quantify the impact of the satellite on the pointing, we then project the satellite location as well as the pointing to a 2D x,y plane.
-In the 2D plane, the field of pointing becomes roughly a circle and the start and end locations of the satellite become two points on the plane.
-Using the shapely python library, we calculate the intersection length of the satellite location points and the circle.
-Therefore, with a given simulated satellite constellation and a schedule of observations, we record the number of satellites that occur in each field of view and measure how long the satellite streaks are.
-Using a realistic estimated satellite streak width of 50 pixels ($\sim10$ arcseconds), we can convert the streak length results into the number of affected image pixels, which allows us to quantitatively measure the efficiency of the dodging algorithm with pixel loss.
+We show three example satellite avoidance maps ready for use by the Rubin scheduler in Figure \ref{fig-simulated-scheduler}.
+It is apparent from Figure \ref{fig-simulated-scheduler} that the simulated OneWeb constellation has more negative area --- regions that should be avoided due to large numbers of illuminated satellites --- than the other two constellations.
+Although OneWeb has fewer satellites than Starlink Gen2, the OneWeb satellites orbit at a higher altitude, meaning that they will be illuminated for a longer portion of the night, and also have a larger impact close to twilight.
+This is why one of the recommendations from \citet{satcon1} is to keep LEO satellites below 600 km altitude.
 
-\section{Results}
-We begin by investigating the efficiency of the added weighting algorithm on dodging satellites in observations.
-As shown in Figure \ref{fig-pixel-loss-weight}, using a conservative estimate of 50 pixels as streak width, we found that higher dodging weights reduces pixels lost to satellite streaks.
-We also found that smaller constellations at lower orbital altitudes (Starlink Gen1 for example) inherently results in less pixel loss per pointing, nearly independent of the dodging weight.
-Figure \ref{fig-pixel-loss-weight} shows that the added weighted dodging algorithm was able to effectively avoid satellite streaks in pointings.
+To investigate whether the scheduler behaves how we expect with the new satellite avoidance strategy, we create a testing function that measures the length of satellite streaks in the simulated field of views.
+To ensure efficiency, only satellites that are above the altitude limit and illuminated by the Sun are considered.
+Satellites below the altitude limit (indicated in the gray region in Figure \ref{fig-simulated-scheduler}) cannot be observed in pointings and are therefore not included.
+For each satellite, we first determine whether it is in the field of view for a given pointing by calculating their distance from the center of the field of view.
+If this distance is less than the radius of the field of view, the satellite has crossed through the pointing.
+To quantify the impact of the satellite on the pointing, we then project both the satellite location and the pointing to a 2D x,y plane.
+In this plane, the field of view is roughly circular and the start and end locations of the satellite crossing are two points on the plane, and we can calculate the total intersection length.
+Therefore, given a simulated satellite constellation and a schedule of observations, we are able to record the number of satellites in each pointing and measure the total streak length.
+Using a typical streak width of 50 pixels ($\sim10$ arcseconds), we finally convert this streak length into a number of affected image pixels, which allows us to quantify the efficiency of the satellite avoidance strategy as a function of pixel loss.
+% TODO: somehow address "typical streak width"
 
-\begin{figure}[ht!]
+We simulate observations for only the first year of the planned ten-year LSST.
+The survey strategy does not significantly change in later years, so we naively extrapolate satellite streak statistics from one year to the entire ten-year survey.
+We acknowledge this does not account for the likely satellite population increase beyond the three simulated constellations.
+We do not consider the effects of satellites launching or de-orbiting, and for simplicity we assume each satellite's orbital parameters are constant.
+This should be an acceptable approximation as long as actual satellite orbital parameters are available $\sim1$\ day in advance so our avoidance basis functions can be constructed.
+If there is no timely information publicly available on LEO satellite constellation orbits, or it is highly inaccurate, the satellite avoidance strategy would be impossible to implement.
+We estimate that satellite orbital solutions correct to within about a degree in space and to within a few minutes in time would be sufficient to effectively avoid some regions of the sky with more satellites in large constellations.
+% TODO: add something about how our results are semi-extrapolateable to larger constellations in similar distributions, which is why considering these three ones is representative of the future (to some degree)
+
+\section{Results}\label{results}
+
+We find that higher dodging weights reduces pixels lost to satellite streaks, and that the dodging algorithm is able to effectively avoid satellite streaks in simulated pointings.
+This is shown in the top two panels of Figure \ref{fig-pixel-loss-weight}, where we adopt 50 pixels as a typical satellite streak width.
+We note satellite streak masks may need to be as much as six times wider than this unless precision modeling and subtraction is implemented \citep{hasan22}.
+We also find that smaller constellations at lower orbital altitudes (Starlink Gen1, for example) inherently cause less pixel loss per pointing, nearly independent of the dodging weight.
+
+\begin{figure*}[ht!]
 \plottwo{plots/streaklen_v_weight.pdf}{plots/streakfrac_v_weight.pdf}
@@
 \caption{Results from simulating the first year of Rubin observations in the presence of different satellite constellations and varying how strongly the scheduler attempts to avoid satellites.
-Top panels show that, as expected, the number of streaks and mean streak length decrease as more emphasis is put on satellite dodging.
-Bottom panels show the resulting trade off, as avoiding satellites forces the scheduler to make longer slews and observe less desirable parts of the sky, resulting in fewer total exposures in the first year and a shallower median coadded depth in $g$.
-xxx--could stack these and share the x-axis. 
+The top panels show that, as expected, the mean streak length per visit (upper left) and fraction of streaked visits (upper right) decrease as more emphasis is put on avoiding satellites.
+The bottom panels show the resulting trade off, as avoiding satellites forces the scheduler to spend more time slewing and observe less desirable parts of the sky.
+This result in fewer total exposures in the first year (lower left) and a shallower median co-added survey depth, shown here in $g$-band (lower right).
 \label{fig-pixel-loss-weight}}
-\end{figure}
+\end{figure*}
 
-We then investigate the relationship between number of exposures taken and the dodging weight.
-As shown in Figure \ref{fig-pixel-loss-weight}, higher dodging weight results in fewer exposures, most likely due to longer slew times.
-With the additional avoidance weighting, the telescope may be prompted to slew to a location other than the desired pointing location and then slew back in order to avoid some satellites, resulting in fewer overall exposures.
-We also found that larger constellations like Starlink Gen2 tend to decrease the number of exposures slightly more than smaller constellations (Starlink Gen1), which is expected.
-More satellites or satellites at higher orbital altitudes result in bigger areas of avoidance on the sky, which results in more slewing to avoid the affected areas, which subsequently reduces exposures.
-\\
+Next, we investigate the relationship between the number of exposures the scheduler is able to complete as a function of the dodging weight.
+As shown in the bottom two panels of Figure \ref{fig-pixel-loss-weight}, higher dodging weight results in fewer visits, most likely due to longer slew times.
+With a higher satellite avoidance weight, the telescope may be prompted to slew to a location other than the desired pointing location and then slew back to avoid some satellites, resulting in fewer overall exposures.
+We also find that a larger constellation (Starlink Gen2) tends to decrease the number of exposures slightly more than a smaller constellation (Starlink Gen1), which is expected.
+More satellites or satellites at higher orbital altitudes result in larger areas of avoidance on the sky, which leads to more slewing required to avoid the affected areas, which subsequently reduces the total number of visits.
 
-We now investigate the tradeoff between survey depth and satellite avoidance.
-One goal of LSST is to collect a large number of exposures of all parts of the southern sky so the exposures can be coadded to reveal faint structures that aren’t visible in individual images.
-Therefore, survey depth is crucial to LSST science goals and the trade off must be evaluated.
-With the satellite avoidance algorithm, the scheduler is prompted to avoid regions with illuminated satellites, which could result in longer slew time or less desirable pointing conditions, which contribute to a loss in survey depth.
-Therefore, we evaluate the tradeoff between survey depth loss and satellite avoidance using the dodging algorithm.\\
+One important LSST survey goal is to collect a large number of exposures of the whole southern sky so these may be co-added to reveal faint structures that are not visible in individual visits.
+Therefore, survey depth is crucial to LSST science, and the trade off between pixel loss from satellite streaks versus survey depth reduction from fewer total visits must be evaluated.
+With the satellite avoidance algorithm, the scheduler is prompted to avoid regions with illuminated satellites, which sometimes results in longer slew times or less desirable pointing conditions and contributes to a loss in survey depth.
+Therefore, Figure \ref{fig-trade-off} explores the trade off between survey depth loss and satellite avoidance using the dodging algorithm.
+As before, we assume a typical satellite streak requires a 50-pixel-wide mask (with a plate scale of 0.2 arcseconds per pixel).
+
+For all three constellations, a minor reduction in mean pixel loss per visit ($0.001\%$ decrease) results in a significant decrease in median co-added depth ($0.02-0.2$ mags).
+Without any satellite avoidance, pixel loss due to satellite streaks only constitutes a small percentage of total pixels in a pointing (about $0.008\%$).
+Given that the percentage is so small, a very minor reduction in mean pixel loss per visit results in significant reduction in image depth.
+While the percentage of lost pixels can be decreased, it reduces co-added depth as it forces the scheduler to make longer slews and observe less desirable parts of the sky (e.g., higher airmass, brighter background).
 
 \begin{figure}[ht!]
-\epsscale{0.7}
 \plotone{plots/streak_length_v_deltam.pdf}
 \epsscale{1}
-\caption{The trade off between pixels lost to streaks and final median coadded depth in $g$\ for the survey.
-We assume streaks require 50 pixel wide masks (with a plate scale of 0.2 arcseconds per pixel).
-Negative changes in coadded depth represent shallower surveys.
-While the percentage of lost pixels can be decreased, it forces the scheduler to make longer slews and observe less desirable parts of the sky (e.g., higher airmass, brighter background).
+\caption{The trade off between the percentage of pixels lost to streaks and final median co-added depth in $g$ band for one year of the LSST, using different satellite avoidance weights.
+A negative change in co-added depth indicates the overall survey is shallower.
+While we can reduce pixel loss due to satellite streaks by half with minimal impact on survey depth, the pixel loss is small to begin with, and this trade off is not worthwhile.
 \label{fig-trade-off}}
 \end{figure}
+% TODO: up the streak width from 50 to 300 pixels and remake Fig 4
 
-As shown in Figure \ref{fig-trade-off} , for all three constellations, minor reduction in mean pixel loss per visit ($0.001\%$ decrease) results in a significant decrease in median coadded depth.
-Currently, pixel loss due to satellite streaks only constitutes a very low percentage of total pixels in a pointing (about $0.008\%$ without any dodging).
-Given that the percentage is currently so small, a very minor reduction in mean pixel loss per visit results in significant reduction in image depth.
+So far, we have primarily considered impacts on the overall LSST.
+However, another important LSST science goal involves using twilight images to search for Near Earth Objects (NEOs).
+These observations must be taken in the direction of the rising or setting sun at high airmass.
+Because a small potential area is targeted, our proposed satellite avoidance scheme is ineffective.
+Figure~\ref{fig:twi_neo} shows how regular survey observations and twilight NEO observations would be affected by satellite constellations using the Rubin scheduler with no satellite avoidance.
+Even in the case of a full Starlink Gen2 constellation, while most exposures would have a streak, only $0.04$\% of twilight NEO observation pixels would be lost, assuming a 50-pixel-wide streak mask.
 
-\begin{figure}
+\begin{figure*}
 \centering
+\epsscale{.8}
+\plottwo{plots/temp/opsim_Count_night_night_lt_366_and_visitExposureTime_gt_20_HEAL_SkyMap.pdf}{plots/regular_curves.pdf}
 \plottwo{plots/temp/opsim_Count_night_night_lt_366_and_visitExposureTime_lt_20_HEAL_SkyMap.pdf}{plots/twilight_curves.pdf}
-    \caption{Impact of streaks on twilight NEO observations that can not easily be shifted to dodge satellites.
-Left panel shows the altitude, azimuth distribution of NEO observations in a recent survey simulation.
-The right panel shows the number of streaks that would result from different satellite constellations as a function of observation altitude.
-The legend includes an estimate of the pixel loss percentage for NEO observations assuming a 50 pixel wide mask is used.
-}
-    \label{fig:twi_neo}
-\end{figure}
+\epsscale{1}
+\caption{Impacts of satellite streaks on simulated LSST observations without any satellite dodging.
+Compared to standard LSST observations (top), twilight NEO observations (bottom) cannot easily be shifted to avoid satellites.
+The left panels show the altitude and azimuth distribution of observations on the sky, and the right panels show how many streaks would result from the three simulated satellite constellations as a function of how high above the horizon the telescope is pointing (observation altitude).
+Even though most twilight NEO observations would contain a satellite streak, pixel loss is still low.
+\label{fig:twi_neo}}
+\end{figure*}
 
-\section{Discussion and Conclusion}
+\section{Discussion}\label{discuss}
 
@@
 This is shown in Figure \ref{fig-pixel-loss-weight} where all three constellations caused less pixel loss with increasing dodging weights.
-However, with the new added priority on dodging, the telescope can be pushed to take an observation path that does not optimize slew time, and subsequently reduces the number of exposures and overall survey depth.
-The tradeoff essentially comes down to the relationship between mean pixel loss reduction and survey depth reduction.
+However, with the new added priority on satellite avoidance, the telescope can be pushed to take an observation path that does not optimize slew time, which subsequently reduces the number of exposures and overall survey depth.
+The trade off essentially comes down to the relationship between mean pixel loss reduction and survey depth reduction.
 If the amount of pixels saved outweighs the amount of survey depth lost, then the algorithm is valuable to be added to the scheduler.
-The tradeoff relationship is evaluated in Figure \ref{fig-trade-off}.
-We found that, for all three constellations, the tradeoff between reduction of pixel loss and loss in coadded image depth is so significant that satellite dodging is overwhelmingly not worth it, provided the simulated constellations we selected are representative of reality.
-Since each pointing contains such a large number of pixels, the number of pixels destroyed by satellite streaks constitutes a very small percentage of the overall image.
-Therefore, while the weighted dodging algorithm is effective at reducing satellite streaks in images, at the current stage, satellite avoidance is mostly unnecessary considering the trade-off between pixel loss and loss in overall survey depth.
+The trade off relationship is evaluated in Figure \ref{fig-trade-off}.
+We find that, for all three constellations, the trade off between reduction of pixel loss and loss in co-added image depth is so significant that satellite avoidance is overwhelmingly not worth it, provided the simulated constellations we selected are representative of reality.
+Since each pointing contains such a large number of pixels, the number of pixels ruined by satellite streaks constitutes a very small percentage of the overall image.
+When evaluating whether to implement a weighted satellite avoidance strategy to effectively reduce satellite streak density, we note it is necessary to evaluate the trade off between pixel loss together with non-linear crosstalk, time-domain glint effects, and any other relevant systematics versus loss in overall survey depth.
 
-xxx--Add short statement about impacts on near-Earth object (NEO) searches, and also a figure that Peter will make.
-XXX--Add in the results from looking at a run with twilight NEO survey observations.
-I think it is a result that will help some folks calm down a bit that we aren't going to miss tons of killer asteroids even with large constellations.
+We note that \citet{lawrence22} stated the majority of LSST images are likely to contain a satellite streak.
+They cited \citet{tyson20}, which further estimated 1\% of LSST twilight pixels would need to be masked given a population of 48,000 LEO satellites.
+The former is only true for twilight observing campaigns, while the latter estimate included higher altitude Starlink orbits that are no longer planned.
+Our study finds that only about 10\% of all LSST images will have a streak from a mega constellations.
+It is true, however, that twilight observing campaigns at high airmass --- like those necessary to perform NEO searches --- will have streaks in the majority of images.
+% TODO: adjust their paragraph with more qualifiers
 
-xxx--Maybe a table for estimated pixel/area losses for different things. Something like narrow streak mask, wide streak mask, chip gaps, and bright/saturated stars.
+Satellite streaks from Starlink and OneWeb as presently designed are not expected to saturate the LSST Camera's CCD detectors.
+In addition, satellites only leave streaks in images when they are both illuminated by the Sun and visible from the observatory --- LEO satellites spend most of the night in Earth's shadow.
+% TODO: add sentence about how larger, higher (in focus) sats may be visible even when they're in Earth's shadow from scattered light
+Thus, even if there are tens of thousands of satellites, the percentage of pixels lost is relatively small.
+For comparison, the Rubin focal plane has a fill factor of 88.0\%, meaning 12\% of potential observations are lost to chip gaps, and losses due to weather can vary by up to 8\%.
+We also expect diffraction spikes from bright stars to render $\sim 1-2$\% of pixels in each exposure unusable.
 
-xxx--Consider saying something like only 10\% of images will have a streak from these three constellations, whereas Lawrence et al.
-2022 says ``a majority''.
-Some of this is probably based on old Starlink info with higher altitudes.
+While the satellite avoidance algorithm we have presented is mostly unnecessary for the three satellite constellations considered, there is still the likelihood for a sharp increases in satellite population in the next $5–10$ years, overlapping the LSST operations period ($2024–2035$).
+With a dramatic increase in satellite population, the ability to avoid satellites might become more relevant.
+In addition, satellites from other operators may be significantly brighter than present-day Starlink and OneWeb satellites, and may saturate the LSST Camera’s detectors or cause overwhelming levels of non-linear crosstalk.
+In this case, avoidance may be required as the number of pixels lost per streak would be much higher.
 
-We find larger constellations will only leave streaks on $\sim10$\% of Rubin images.
-This is a much lower value than previous studies \citep{lawrence22, tyson20}.
-This is partly due to the latest Starlink plans not including higher altitude orbits that were included in the \citet{tyson} study.
-% And partly due to not conflating twilight observations with the whole survey.
+Related to this, one future work direction involves adding brightness weighting to the satellite avoidance algorithm.
+The idea is to only avoid satellites brighter than a certain brightness threshold.
+This could potentially reduce the region of avoidance, therefore reducing the loss in co-added depth --- it could reduce the trade off between pixel loss and loss in survey depth, making this preliminary work directly relevant for a successful LSST.
+It may be possible to compute optimal starting locations for a series of observations based on satellite forecasts to further optimize satellite avoidance.
+Finally, since faint trail detection and masking is not perfect, no satellite avoidance strategy will effectively mitigate faint glints and the resulting bogus alerts.
 
-XXX--While satellite streaks are poised to be a nuisance to Rubin observations, it is important to remember 1) They do not saturate detectors, 2) Do not leave streaks when in earth's shadow, 3)
-
-\section{Future Work}
-While the satellite dodging algorithm is mostly unnecessary for the three satellite constellations we considered, many astronomers estimate that there will be sharp increases in satellite population in the next 5–10 years, which notably overlaps the LSST operations period (2024–2035).
-With the dramatic increase in satellite population, the dodging algorithm might become more relevant.
-In addition, satellites from other operators may be significantly brighter than present-day Starlink and OneWeb satellites, and may saturate the LSST Camera’s detectors.
-In this scenario, dodging may be required to salvage some science, even if it means losing out on survey depth.
-Related to this, one future work direction involves adding in brightness weighting to the dodging algorithm.
-The idea is to only avoid satellites brighter than a certain brightness threshold.
-This could potentially reduce the region of avoidance, therefore reducing the loss in coadded depth — it could reduce the tradeoff between pixel loss and loss in survey depth, making this preliminary work directly relevant for a successful LSST.
-Finally, it may be possible to compute optimal starting locations for observation schedules based on predicted satellite maps in order to optimize satellite avoidance.
-
-\software{Astropy \citep{astropy2013, astropy2018}, 
+\software{Astropy \citep{astropy2013, astropy2018, astropy2022}, 
           Healpy and HEALpix\footnote{\url{http://healpix.sourceforge.net}} \citep{healpix2005, Zonca2019},
@@
           Shapely \citep{shapely2007}}
-          % todo: add citation to newest 2022 astropy paper
 
@@
 
+\begin{acknowledgments}
+JAH acknowledges support from the Computing Research Association-Widening Participation Distributed Research Experiences for Undergraduates (DREU) program.
+JAH and MLR are grateful for LSST Corporation travel support for JAH to attend the 2022 Rubin Project and Community Workshop and present a poster.
+MLR acknowledges support from Bob Blum, Leanne Guy, and all of Rubin Operations to spend a fraction of her time on satellite mitigation work; this study would not have been possible without formal recognition that the proliferation of bright commercial LEO satellites poses a threat to LSST science.
+The authors all wish to thank Tony Tyson for valuable discussions and feedback that helped place this work in context.
+This work was facilitated through the use of advanced computational, storage, and networking infrastructure provided by the Hyak supercomputer system at the University of Washington.
+\end{acknowledgments}
+
 \end{document}

### 2022-11-01 ms.tex
Updates from Overleaf

@@
 Rubin Observatory Legacy Survey of Space and Time (LSST).
-To do this, we construct and simulate the motions of planned Starlink and OneWeb satellites, test how effectively the Rubin scheduler algorithm can avoid them, and assess how the overall survey is affected.
+To do this, we construct and simulate the motions of a total of 40,678 planned Starlink and OneWeb satellites, test how effectively the Rubin scheduler algorithm can avoid them, and assess how the overall survey is affected.
 Given a reasonably accurate satellite orbit forecast, we find it is possible to adjust the scheduler algorithm to effectively avoid some satellites.
-Doing so would reduce pixel loss by a factor of two with minimal impact to the survey depth.
-However, the small fraction of pixels lost to streaks from Starlink and OneWeb may render this mitigation strategy unnecessary.
-With both of these satellite constellations fully deployed, and with no satellite avoidance strategy, the LSST would lose $\sim$0.02\% of science pixels to satellite streaks, assuming a typical streak width of 50 pixels, while high-airmass twilight observations would lose $\sim$0.06\% of science pixels.
-However, it is important to note that false alerts from glints and low surface brightness residuals can have more science impact than pixel loss.
+Overall, reallocating 10\% of LSST observing time to satellite avoidance reduces the fraction of LSST visits with streaks by a factor of two, from 10\% to 5\%.
+The small fraction of pixels lost to streaks from Starlink and OneWeb may render this mitigation strategy unnecessary, however, it is important to note that false alerts from glints and low surface brightness residuals can have more science impact than the number of pixels lost.
 An increase in the number or brightness of satellites during Rubin Operations, together with science impacts not well-represented by pixel loss, may make implementing this satellite avoidance strategy worthwhile.
@@
 
-% TODO: change pixel loss takeaway in abstract to the 10% / factor of 2 thing
+%With both of these satellite constellations fully deployed, and with no satellite avoidance strategy, the LSST would lose $\sim$0.12\% of science pixels to satellite streaks, assuming a typical streak width of 300 pixels, while high-airmass twilight observations would lose $\sim$0.36\% of science pixels.
 
@@
 To accomplish this, the LSST will employ a scheduling algorithm that uses a modified Markov Decision Process which can generate lists of desirable observations in real time \citep{naghib19}.
-The Rubin scheduler balances the desire to minimize slew time, optimize signal to noise in each image, and to maintain a uniform survey footprint.
+The Rubin scheduler balances the desire to minimize slew time, optimize signal to noise in each image, and to maintain uniform survey co-added depth.
 
@@
 \caption{Three simulated satellite constellations, one per column.
+Starlink Gen1 is 4,408 satellites, Starlink Gen2 is 29,988 satellites, and OneWeb is 6,372 satellites, for a grand total of 40,768.
 The top row shows the 3D distribution of each constellation around Earth.
-The middle row shows Hammer projections of the altitude and azimuth positions of each constellation as seen from Rubin Observatory on October 1, 2023 during twilight (Sun altitude $-18$ degrees).
+The middle row shows instantaneous Hammer projections of the altitude and azimuth positions of each constellation as seen from Rubin Observatory on October 1, 2023 during twilight (Sun altitude $-18$ degrees).
 Blue points are satellites illuminated by the Sun at this time, red points are satellites not illuminated by the Sun, and black points are satellites that are both illuminated and above the Rubin 20 degree altitude pointing limit.
-The bottom row is the same Hammer projections 6 hours later in the middle of the night (Sun altitude is $-50$ degrees).
-Because the Starlink satellites orbit at 550 km, none are illuminated in the middle of the night at this time of year.
+The bottom row is the same Hammer projections six hours later in the middle of the night (Sun altitude $-50$ degrees).
+Because Starlink satellites orbit at 550 km, none are illuminated in the middle of the night at this time of year.
 The OneWeb constellation at 1200 km has only a single illuminated satellite above the Rubin altitude limit at this particular time.
+\vspace{4em}
 \label{fig-simulated-constellations}
@@
 \end{figure*}
-% TODO: add number of sats in figure caption, make it clear the bottom 2 rows are instantaneous/snapshot (not integration)
 
@@
 
-The baseline Rubin observing strategy uses three primary basis functions which reward (1) minimizing slewtime, (2) maximizing the depth of images (e.g., by avoiding the Moon and high airmass), and (3) maintaining a uniform depth survey footprint.
+The baseline Rubin observing strategy uses three primary basis functions which reward (1) minimizing slewtime, (2) maximizing the depth of images (e.g., by avoiding the Moon and high airmass), and (3) maintaining a uniform depth survey co-added depth.
 To this, we add a fourth basis function which rewards avoiding areas of the sky which will have high concentrations of illuminated satellites.
@@
 Darker regions have more illuminated satellites and therefore more negative weighting.
-By varying the weight placed on these maps, the scheduler will more actively avoid regions of the sky where satellites could streak images.
+By varying the ``dodging weight'' placed on these maps, the scheduler will more actively avoid regions of the sky where satellites could streak images.
+\vspace{3em}
 \label{fig-simulated-scheduler}
@@
 To ensure efficiency, only satellites that are above the altitude limit and illuminated by the Sun are considered.
-Satellites below the altitude limit (indicated in the gray region in Figure \ref{fig-simulated-scheduler}) cannot be observed in pointings and are therefore not included.
+Satellites below the altitude limit (indicated in the gray region in Figure \ref{fig-simulated-scheduler}) cannot be reached by the Simonyi Survey Telescope and are therefore not included.
 For each satellite, we first determine whether it is in the field of view for a given pointing by calculating their distance from the center of the field of view.
@@
 Therefore, given a simulated satellite constellation and a schedule of observations, we are able to record the number of satellites in each pointing and measure the total streak length.
-Using a typical streak width of 50 pixels ($\sim10$ arcseconds), we finally convert this streak length into a number of affected image pixels, which allows us to quantify the efficiency of the satellite avoidance strategy as a function of pixel loss.
-% TODO: somehow address "typical streak width"
+Using a conservative streak width of 300 pixels \citep[$\sim1$ arcminute per][]{hasan22}, we finally convert this streak length into a number of affected image pixels, which allows us to quantify the efficiency of the satellite avoidance strategy as a function of pixel loss.
 
@@
 The survey strategy does not significantly change in later years, so we naively extrapolate satellite streak statistics from one year to the entire ten-year survey.
-We acknowledge this does not account for the likely satellite population increase beyond the three simulated constellations.
+We acknowledge this does not account for the likely satellite population increase beyond the three simulated constellations; however, our results should scale linearly to larger future constellations in similar orbital distributions.
 We do not consider the effects of satellites launching or de-orbiting, and for simplicity we assume each satellite's orbital parameters are constant.
@@
 We estimate that satellite orbital solutions correct to within about a degree in space and to within a few minutes in time would be sufficient to effectively avoid some regions of the sky with more satellites in large constellations.
-% TODO: add something about how our results are semi-extrapolateable to larger constellations in similar distributions, which is why considering these three ones is representative of the future (to some degree)
 
@@
 We find that higher dodging weights reduces pixels lost to satellite streaks, and that the dodging algorithm is able to effectively avoid satellite streaks in simulated pointings.
-This is shown in the top two panels of Figure \ref{fig-pixel-loss-weight}, where we adopt 50 pixels as a typical satellite streak width.
-We note satellite streak masks may need to be as much as six times wider than this unless precision modeling and subtraction is implemented \citep{hasan22}.
+This is shown in the top two panels of Figure \ref{fig-pixel-loss-weight}.
 We also find that smaller constellations at lower orbital altitudes (Starlink Gen1, for example) inherently cause less pixel loss per pointing, nearly independent of the dodging weight.
@@
 \plottwo{plots/nexp_v_weight.pdf}{plots/deltam_v_weight.pdf}
-\caption{Results from simulating the first year of Rubin observations in the presence of different satellite constellations and varying how strongly the scheduler attempts to avoid satellites.
+\caption{To reduce the fraction of visits with streaks by a factor of two, satellite avoidance will require 10\% of total observing time.
+%Results from simulating the first year of Rubin observations in the presence of different satellite constellations and varying how strongly the scheduler attempts to avoid satellites.
 The top panels show that, as expected, the mean streak length per visit (upper left) and fraction of streaked visits (upper right) decrease as more emphasis is put on avoiding satellites.
@@
 Therefore, Figure \ref{fig-trade-off} explores the trade off between survey depth loss and satellite avoidance using the dodging algorithm.
-As before, we assume a typical satellite streak requires a 50-pixel-wide mask (with a plate scale of 0.2 arcseconds per pixel).
+Following \citet{hasan22}, we assume a typical satellite streak requires a 300-pixel-wide mask (equivalent to 1 arcminute given the plate scale of 0.2 arcseconds per pixel).
 
-For all three constellations, a minor reduction in mean pixel loss per visit ($0.001\%$ decrease) results in a significant decrease in median co-added depth ($0.02-0.2$ mags).
-Without any satellite avoidance, pixel loss due to satellite streaks only constitutes a small percentage of total pixels in a pointing (about $0.008\%$).
+For all three constellations, a minor reduction in mean pixel loss per visit ($0.006\%$ decrease) results in a significant decrease in median co-added depth ($0.02-0.2$ mags).
+Without any satellite avoidance, pixel loss due to satellite streaks only constitutes a small percentage of total pixels in a pointing (about $0.048\%$).
 Given that the percentage is so small, a very minor reduction in mean pixel loss per visit results in significant reduction in image depth.
@@
 \epsscale{1}
-\caption{The trade off between the percentage of pixels lost to streaks and final median co-added depth in $g$ band for one year of the LSST, using different satellite avoidance weights.
+\caption{The trade off between the percentage of pixels lost to streaks and final median co-added depth in $g$ band for one year of the LSST, using different satellite avoidance weights and assuming streak masks are 300 pixels wide \citep{hasan22}.
 A negative change in co-added depth indicates the overall survey is shallower.
-While we can reduce pixel loss due to satellite streaks by half with minimal impact on survey depth, the pixel loss is small to begin with, and this trade off is not worthwhile.
+We can reduce pixel loss due to satellite streaks by half with minimal impact on survey depth.
 \label{fig-trade-off}}
 \end{figure}
-% TODO: up the streak width from 50 to 300 pixels and remake Fig 4
 
@@
 Figure~\ref{fig:twi_neo} shows how regular survey observations and twilight NEO observations would be affected by satellite constellations using the Rubin scheduler with no satellite avoidance.
-Even in the case of a full Starlink Gen2 constellation, while most exposures would have a streak, only $0.04$\% of twilight NEO observation pixels would be lost, assuming a 50-pixel-wide streak mask.
+Even in the case of a full Starlink Gen2 constellation, while most exposures would have a streak, only $0.24$\% of twilight NEO observation pixels would be lost, assuming a 300-pixel-wide streak mask.
 
@@
 \centering
-\epsscale{.8}
+\epsscale{.85}
 \plottwo{plots/temp/opsim_Count_night_night_lt_366_and_visitExposureTime_gt_20_HEAL_SkyMap.pdf}{plots/regular_curves.pdf}
@@
 \epsscale{1}
-\caption{Impacts of satellite streaks on simulated LSST observations without any satellite dodging.
+\caption{Impacts of satellite streaks on simulated LSST observations without any satellite avoidance.
 Compared to standard LSST observations (top), twilight NEO observations (bottom) cannot easily be shifted to avoid satellites.
 The left panels show the altitude and azimuth distribution of observations on the sky, and the right panels show how many streaks would result from the three simulated satellite constellations as a function of how high above the horizon the telescope is pointing (observation altitude).
-Even though most twilight NEO observations would contain a satellite streak, pixel loss is still low.
+While most twilight NEO observations would contain a satellite streak, pixel loss is still low.
 \label{fig:twi_neo}}
@@
 They cited \citet{tyson20}, which further estimated 1\% of LSST twilight pixels would need to be masked given a population of 48,000 LEO satellites.
-The former is only true for twilight observing campaigns, while the latter estimate included higher altitude Starlink orbits that are no longer planned.
-Our study finds that only about 10\% of all LSST images will have a streak from a mega constellations.
+The former is only true for twilight observing campaigns, while the latter estimate included some higher-altitude Starlink orbits that are no longer planned.
+Our study finds that only about 10\% of all LSST images will have a streak from the three simulated constellations (Starlink Gen1, Starlink Gen2, and OneWeb, totaling 40,768 satellites as currently planned).
 It is true, however, that twilight observing campaigns at high airmass --- like those necessary to perform NEO searches --- will have streaks in the majority of images.
-% TODO: adjust their paragraph with more qualifiers
+Extrapolating these results to a possible LEO satellite population in the hundreds of thousands circa 2030 is beyond the scope of this work.
 
 Satellite streaks from Starlink and OneWeb as presently designed are not expected to saturate the LSST Camera's CCD detectors.
-In addition, satellites only leave streaks in images when they are both illuminated by the Sun and visible from the observatory --- LEO satellites spend most of the night in Earth's shadow.
-% TODO: add sentence about how larger, higher (in focus) sats may be visible even when they're in Earth's shadow from scattered light
-Thus, even if there are tens of thousands of satellites, the percentage of pixels lost is relatively small.
-For comparison, the Rubin focal plane has a fill factor of 88.0\%, meaning 12\% of potential observations are lost to chip gaps, and losses due to weather can vary by up to 8\%.
+Satellites typically only leave streaks in images when they are both illuminated by the Sun and visible from the observatory.
+LEO satellites spend most of the night in Earth's shadow.
+However, it is possible for larger satellites in higher orbits, where they appear more in focus --- like OneWeb at 1200 km --- to leave faint scattered light artifacts that may be visible to the LSST Camera.
+This possibility notwithstanding, even with all three simulated constellations in orbit, the percentage of pixels lost is relatively small.
+For comparison, the Rubin focal plane has a fill factor of 88.0\%\footnote{\url{https://smtn-004.lsst.io}}, meaning 12\% of potential observations are lost to chip gaps, and losses due to weather can vary by up to 8\%.
 We also expect diffraction spikes from bright stars to render $\sim 1-2$\% of pixels in each exposure unusable.
@@
 \begin{acknowledgments}
-JAH acknowledges support from the Computing Research Association-Widening Participation Distributed Research Experiences for Undergraduates (DREU) program.
+JAH acknowledges support from the Computing Research Association-Widening Participation (CRA-WP) Distributed Research Experiences for Undergraduates (DREU) program.
 JAH and MLR are grateful for LSST Corporation travel support for JAH to attend the 2022 Rubin Project and Community Workshop and present a poster.

### 2022-11-04 ms.tex

@@
 Rubin Observatory Legacy Survey of Space and Time (LSST).
-To do this, we construct and simulate the motions of a total of 40,678 planned Starlink and OneWeb satellites, test how effectively the Rubin scheduler algorithm can avoid them, and assess how the overall survey is affected.
+We simulate the orbits of currently planned Starlink and OneWeb constellations ($\sim$40,000 satellites) to test how effectively an upgraded Rubin scheduler algorithm can avoid them, and assess how the overall survey is affected.
 Given a reasonably accurate satellite orbit forecast, we find it is possible to adjust the scheduler algorithm to effectively avoid some satellites.
-Overall, reallocating 10\% of LSST observing time to satellite avoidance reduces the fraction of LSST visits with streaks by a factor of two, from 10\% to 5\%.
-The small fraction of pixels lost to streaks from Starlink and OneWeb may render this mitigation strategy unnecessary, however, it is important to note that false alerts from glints and low surface brightness residuals can have more science impact than the number of pixels lost.
-An increase in the number or brightness of satellites during Rubin Operations, together with science impacts not well-represented by pixel loss, may make implementing this satellite avoidance strategy worthwhile.
+Overall, sacrificing 10\% of LSST observing time to avoid satellites reduces the fraction of LSST visits with streaks by a factor of two.
+Whether such mitigation will be required depends on the overall impact of streaks on science which is not well quantified yet due to a lack of adequate information about satellite brightness distribution and the impact of glints and low surface brightness residuals on alert purity
+and systematic errors in cosmological parameter estimation.
+A significant increase in the number of satellites or their brightness during Rubin Operations may make implementing this satellite avoidance strategy worthwhile.
 \end{abstract}
@@
 Rubin Observatory's LSST is a ten-year astronomical imaging survey that will begin in 2024 from a new telescope under construction in Chile.
-Instead of soliciting individual requests for what the telescope should observe, the LSST will uniformly survey the sky every few nights using six color filters to essentially create a decade-long high-resolution survey of the entire southern sky, and share massive quantities of data products with the astronomy community \citep{overview}.
+Instead of soliciting individual requests for what the telescope should observe, the LSST will uniformly survey the sky every few nights using six color filters to create a decade-long high-resolution survey of the entire southern sky, and share massive quantities of data products with the astronomy community \citep{overview}.
 To accomplish this, the LSST will employ a scheduling algorithm that uses a modified Markov Decision Process which can generate lists of desirable observations in real time \citep{naghib19}.
-The Rubin scheduler balances the desire to minimize slew time, optimize signal to noise in each image, and to maintain uniform survey co-added depth.
+The Rubin scheduler balances the desire to minimize slew time, optimize signal to noise in individual images, and to maintain survey footprint uniformity.
 
@@
 LEO satellites are visible from Earth because they reflect sunlight, especially during twilight.
-As the illuminated satellites move across the field of view of an astronomical pointing, they leave a streak in the image.
+As the sun-illuminated satellites move across the field of view of an astronomical pointing, they leave a streak in the image.
 While the flux from satellite streaks can in many cases be identified and removed, the resulting pixels have much lower signal-to-noise.
@@
 \citet{tyson20} used a very simple algorithm to see if the LSST could avoid imaging satellite streaks.
-They concluded that attempting to naively ``dodge'' of order 48,000 LEO satellites is useless because it is operationally inefficient.
+They concluded that attempting to naively dodge of order 48,000 LEO satellites is useless because it is operationally inefficient.
 In this paper, rather than try to avoid all satellite streaks, we incorporate satellite avoidance as a component of the Rubin scheduler's Markov Decision Process.
-This allows us to avoid a significant fraction of satellite streaks and investigate what level of avoidance might be acceptable because it does not drastically alter the overall performance of the LSST.
+This allows us to avoid a significant fraction of satellite streaks and investigate what level of avoidance might be acceptable because it does not drastically impact the overall performance of the LSST.
 
@@
 This is discussed in more detail in \citet{tyson20} and on the Rubin Observatory LSST Project website\footnote{\url{https://ls.st/satcon}}.
-Astronomers have also developed algorithms for masking satellite trails in images, but covering the outer rim of the trails without losing extra pixels remains a challenge \citep{hasan22}.
+Astronomers have also developed algorithms for masking satellite trails in images, but covering the outer wings of the trails without losing extra pixels remains a challenge \citep{hasan22}.
 The rapid increase in population of LEO satellites threatens to compromise the quality and scientific value of LSST images and also requires extra human and computer resources to effectively mask trails.
@@
 The top row shows the 3D distribution of each constellation around Earth.
-The middle row shows instantaneous Hammer projections of the altitude and azimuth positions of each constellation as seen from Rubin Observatory on October 1, 2023 during twilight (Sun altitude $-18$ degrees).
+The middle row shows an instantaneous Hammer projection of the altitude and azimuth positions of each constellation as seen from Rubin Observatory on October 1, 2023 during twilight (Sun altitude $-18$ degrees).
 Blue points are satellites illuminated by the Sun at this time, red points are satellites not illuminated by the Sun, and black points are satellites that are both illuminated and above the Rubin 20 degree altitude pointing limit.
@@
 
-The baseline Rubin observing strategy uses three primary basis functions which reward (1) minimizing slewtime, (2) maximizing the depth of images (e.g., by avoiding the Moon and high airmass), and (3) maintaining a uniform depth survey co-added depth.
-To this, we add a fourth basis function which rewards avoiding areas of the sky which will have high concentrations of illuminated satellites.
-Figure~\ref{fig-simulated-scheduler} shows an example of this new basis function for the three simulated satellite constellations at two different Sun altitudes.
+The baseline Rubin observing strategy uses three primary basis functions which reward (1) minimizing slewtime, (2) maximizing the depth of images (e.g., by avoiding the Moon and high airmass), and (3) maintaining a uniform survey footprint.
+To this, we add a fourth basis function which penalizes observing areas of the sky which will have high concentrations of illuminated satellites.
+Figure~\ref{fig-simulated-scheduler} shows an example of this new basis function for the three simulated satellite constellations.
 The positions of the illuminated satellites are computed in 10-second intervals and marked on the sky.
@@
 Thus our modified scheduler with a satellite avoidance strategy does not try to avoid individual satellite streaks, but rather has a parameterized method for avoiding regions of the sky where satellite streaks are more likely.
-This has an additional benefit of not requiring high precision satellite orbit forecasts.
+This has the additional benefit of not requiring high precision satellite orbit forecasts.
 
@@
 Each is for a twilight observation period of 90 minutes (beginning after sunset with a Sun altitude of $-17.1$ degrees).
-The maps are rotated so zenith is in the center of the image.
+The map projections are rotated so zenith is in the center of the image.
 Darker regions have more illuminated satellites and therefore more negative weighting.
-By varying the ``dodging weight'' placed on these maps, the scheduler will more actively avoid regions of the sky where satellites could streak images.
-\vspace{3em}
+By varying the dodging weight placed on these maps, the scheduler will more actively avoid regions of the sky where satellites could streak images.
+%\vspace{3em}
 \label{fig-simulated-scheduler}
@@
 
-To investigate whether the scheduler behaves how we expect with the new satellite avoidance strategy, we create a testing function that measures the length of satellite streaks in the simulated field of views.
+To investigate whether the scheduler behaves how we expect with the new satellite avoidance strategy, we create a testing function that measures the length of satellite streaks in the simulated fields of view.
 To ensure efficiency, only satellites that are above the altitude limit and illuminated by the Sun are considered.
@@
 Therefore, given a simulated satellite constellation and a schedule of observations, we are able to record the number of satellites in each pointing and measure the total streak length.
-Using a conservative streak width of 300 pixels \citep[$\sim1$ arcminute per][]{hasan22}, we finally convert this streak length into a number of affected image pixels, which allows us to quantify the efficiency of the satellite avoidance strategy as a function of pixel loss.
+We assume that
+the impact of streaks on science is proportional to their total length, which allows us to quantify the efficiency of the satellite avoidance strategy.
+
+To estimate the fraction of pixels affected by streaks, following \citet{hasan22} we adopt a fiducial width of 300 pixels (equivalent to 1 arcminute given the plate scale of 0.2 arcseconds per pixel).
+Assuming a length of 15 CCDs (with a CCD side of 4096 pixels), a single streak
+corresponds to 0.6\% of all the pixels in the focal plane. On the other hand, if a streak
+is so bright that entire CCDs are rendered scientifically useless, a single streak
+would wipe out 8\% of all pixels in the focal plane.
+At this time, the overall impact of streaks on science cannot be well quantified due to a lack of adequate information about satellite brightness distribution and the impact of glints and low surface brightness residuals on alert purity and systematic errors in cosmological parameter estimation.
+That is, not only that the fraction of affected pixels is an imperfect metric for 
+assessing the impact of streaks on science, but even this fraction cannot be yet
+well determined in an absolute sense because the effective streak width depends on unknown satellite brightness.
 
 We simulate observations for only the first year of the planned ten-year LSST.
-The survey strategy does not significantly change in later years, so we naively extrapolate satellite streak statistics from one year to the entire ten-year survey.
+The survey strategy does not significantly change in later years.
 We acknowledge this does not account for the likely satellite population increase beyond the three simulated constellations; however, our results should scale linearly to larger future constellations in similar orbital distributions.
@@
 
-We find that higher dodging weights reduces pixels lost to satellite streaks, and that the dodging algorithm is able to effectively avoid satellite streaks in simulated pointings.
+We find that higher dodging weights reduces pixels lost to satellite streaks, and that the satellite avoidance algorithm is able to effectively avoid satellite streaks in simulated pointings.
 This is shown in the top two panels of Figure \ref{fig-pixel-loss-weight}.
@@
 \plottwo{plots/nexp_v_weight.pdf}{plots/deltam_v_weight.pdf}
-\caption{To reduce the fraction of visits with streaks by a factor of two, satellite avoidance will require 10\% of total observing time.
-%Results from simulating the first year of Rubin observations in the presence of different satellite constellations and varying how strongly the scheduler attempts to avoid satellites.
-The top panels show that, as expected, the mean streak length per visit (upper left) and fraction of streaked visits (upper right) decrease as more emphasis is put on avoiding satellites.
-The bottom panels show the resulting trade off, as avoiding satellites forces the scheduler to spend more time slewing and observe less desirable parts of the sky.
-This result in fewer total exposures in the first year (lower left) and a shallower median co-added survey depth, shown here in $g$-band (lower right).
+\epsscale{1.2}
+\caption{Illustration of changes in the mean streak length in all visits including those with no streak (top left), the fraction of visits with streaks
+(top right), the number of acquired visits in year 1 (bottom left) and coadded depth in the $g$ band (bottom right) as a function of dodging basis function weight (starting with essentially no dodging on the left).
+Note that to reduce the fraction of visits with streaks by about a factor of two, satellite avoidance will require 10\% of total observing time.
+\vspace{3em}
 \label{fig-pixel-loss-weight}}
@@
 As shown in the bottom two panels of Figure \ref{fig-pixel-loss-weight}, higher dodging weight results in fewer visits, most likely due to longer slew times.
-With a higher satellite avoidance weight, the telescope may be prompted to slew to a location other than the desired pointing location and then slew back to avoid some satellites, resulting in fewer overall exposures.
+With a higher dodging weight, the telescope may be prompted to slew to a location other than the most desirable nearby pointings, resulting in fewer overall exposures.
 We also find that a larger constellation (Starlink Gen2) tends to decrease the number of exposures slightly more than a smaller constellation (Starlink Gen1), which is expected.
 More satellites or satellites at higher orbital altitudes result in larger areas of avoidance on the sky, which leads to more slewing required to avoid the affected areas, which subsequently reduces the total number of visits.
+In addition to forcing longer slews, avoiding satellite dense areas pushes the scheduler to observe pointings with lower signal to noise (e.g., higher airmass, brighter sky background areas) than it normally would.
 
 One important LSST survey goal is to collect a large number of exposures of the whole southern sky so these may be co-added to reveal faint structures that are not visible in individual visits.
-Therefore, survey depth is crucial to LSST science, and the trade off between pixel loss from satellite streaks versus survey depth reduction from fewer total visits must be evaluated.
+Therefore, survey depth is crucial to LSST science, and the trade-off between pixel loss from satellite streaks versus survey depth reduction from fewer total visits must be evaluated.
 With the satellite avoidance algorithm, the scheduler is prompted to avoid regions with illuminated satellites, which sometimes results in longer slew times or less desirable pointing conditions and contributes to a loss in survey depth.
-Therefore, Figure \ref{fig-trade-off} explores the trade off between survey depth loss and satellite avoidance using the dodging algorithm.
-Following \citet{hasan22}, we assume a typical satellite streak requires a 300-pixel-wide mask (equivalent to 1 arcminute given the plate scale of 0.2 arcseconds per pixel).
-
-For all three constellations, a minor reduction in mean pixel loss per visit ($0.006\%$ decrease) results in a significant decrease in median co-added depth ($0.02-0.2$ mags).
-Without any satellite avoidance, pixel loss due to satellite streaks only constitutes a small percentage of total pixels in a pointing (about $0.048\%$).
-Given that the percentage is so small, a very minor reduction in mean pixel loss per visit results in significant reduction in image depth.
-While the percentage of lost pixels can be decreased, it reduces co-added depth as it forces the scheduler to make longer slews and observe less desirable parts of the sky (e.g., higher airmass, brighter background).
+Therefore, Figure \ref{fig-trade-off} explores the trade-off between survey depth loss and satellite avoidance.
+As evident
+from the figure, the fraction of LSST visits with streaks can be decreased by a factor of two with 
+an investment of 10\% of LSST observing time, corresponding to a loss of coadded depth of 0.05 mag.
 
@@
 \plotone{plots/streak_length_v_deltam.pdf}
-\epsscale{1}
-\caption{The trade off between the percentage of pixels lost to streaks and final median co-added depth in $g$ band for one year of the LSST, using different satellite avoidance weights and assuming streak masks are 300 pixels wide \citep{hasan22}.
-A negative change in co-added depth indicates the overall survey is shallower.
-We can reduce pixel loss due to satellite streaks by half with minimal impact on survey depth.
+\epsscale{1.5}
+\caption{The trade-off between the mean streak length and final median co-added depth in $g$ band for one year of the LSST (controlled by dodging basis function weight).
+A negative change in co-added depth indicates the survey is shallower.
+%\vspace{3em}
 \label{fig-trade-off}}
@@
 So far, we have primarily considered impacts on the overall LSST.
-However, another important LSST science goal involves using twilight images to search for Near Earth Objects (NEOs).
+However, another important LSST science goal involves using twilight images to search for Near Earth Objects (NEOs) and solar system objects interior to the orbits of Earth and Venus.
 These observations must be taken in the direction of the rising or setting sun at high airmass.
@@
 Figure~\ref{fig:twi_neo} shows how regular survey observations and twilight NEO observations would be affected by satellite constellations using the Rubin scheduler with no satellite avoidance.
-Even in the case of a full Starlink Gen2 constellation, while most exposures would have a streak, only $0.24$\% of twilight NEO observation pixels would be lost, assuming a 300-pixel-wide streak mask.
+While the majority of twilight NEO observations could include a satellite streak with a Starlink Gen2 size constellation, we find this would only result in a $\sim$0.36\% loss of science pixels.
 
@@
 The left panels show the altitude and azimuth distribution of observations on the sky, and the right panels show how many streaks would result from the three simulated satellite constellations as a function of how high above the horizon the telescope is pointing (observation altitude).
-While most twilight NEO observations would contain a satellite streak, pixel loss is still low.
+Note that  most twilight NEO observations would contain a satellite streak.
+%\vspace{3em}
 \label{fig:twi_neo}}
@@
 
-We have demonstrated that adding a weighted term in the scheduler algorithm for illuminated satellites can effectively reduce the amount of satellite streaks in observations, and subsequently reduce mean pixel loss per visit.
-This is shown in Figure \ref{fig-pixel-loss-weight} where all three constellations caused less pixel loss with increasing dodging weights.
+We have demonstrated that adding a weighted term in the scheduler algorithm for illuminated satellites can effectively reduce the amount of satellite streaks in observations, and subsequently reduce mean pixel loss per visit and other impacts on science, as illustrated in Figure \ref{fig-pixel-loss-weight}.
 However, with the new added priority on satellite avoidance, the telescope can be pushed to take an observation path that does not optimize slew time, which subsequently reduces the number of exposures and overall survey depth.
-The trade off essentially comes down to the relationship between mean pixel loss reduction and survey depth reduction.
-If the amount of pixels saved outweighs the amount of survey depth lost, then the algorithm is valuable to be added to the scheduler.
-The trade off relationship is evaluated in Figure \ref{fig-trade-off}.
-We find that, for all three constellations, the trade off between reduction of pixel loss and loss in co-added image depth is so significant that satellite avoidance is overwhelmingly not worth it, provided the simulated constellations we selected are representative of reality.
-Since each pointing contains such a large number of pixels, the number of pixels ruined by satellite streaks constitutes a very small percentage of the overall image.
-When evaluating whether to implement a weighted satellite avoidance strategy to effectively reduce satellite streak density, we note it is necessary to evaluate the trade off between pixel loss together with non-linear crosstalk, time-domain glint effects, and any other relevant systematics versus loss in overall survey depth.
+The trade-off essentially comes down to the relationship between total streak length reduction and survey depth reduction.
+The final decision will depend on the overall impact of streaks on science which is not well quantified yet due to a lack of adequate information about satellite brightness distribution and the impact of glints and low surface brightness residuals on alert purity and systematic errors in cosmological parameter estimation.
+In other words,
+when evaluating whether to implement a weighted satellite avoidance strategy to effectively reduce satellite streak density, it is necessary to evaluate the trade-off between pixel loss together with non-linear crosstalk, time-domain glint effects, and any other relevant systematics versus loss in observing time.
 
-We note that \citet{lawrence22} stated the majority of LSST images are likely to contain a satellite streak.
-They cited \citet{tyson20}, which further estimated 1\% of LSST twilight pixels would need to be masked given a population of 48,000 LEO satellites.
-The former is only true for twilight observing campaigns, while the latter estimate included some higher-altitude Starlink orbits that are no longer planned.
-Our study finds that only about 10\% of all LSST images will have a streak from the three simulated constellations (Starlink Gen1, Starlink Gen2, and OneWeb, totaling 40,768 satellites as currently planned).
+We note that earlier publications \citep{lawrence22,tyson20}, stated that the majority of LSST images are likely to contain a satellite streak.
+They also included some higher-altitude Starlink orbits that are no longer planned.
+Our study finds that about 10\% of all LSST images will have a streak from the three simulated constellations (Starlink Gen1, Starlink Gen2, and OneWeb, totaling 40,768 satellites as currently planned).
 It is true, however, that twilight observing campaigns at high airmass --- like those necessary to perform NEO searches --- will have streaks in the majority of images.
-Extrapolating these results to a possible LEO satellite population in the hundreds of thousands circa 2030 is beyond the scope of this work.
 
+There is a concerning possibility of sharp increases in satellite population in the next $5–10$ years, overlapping the LSST operations period ($2024–2035$).
+With a dramatic increase in satellite population, the ability to avoid satellites might become more relevant.
+It is possible to linearly extrapolate our results to consider a possible LEO satellite population in the hundreds of thousands circa 2030, assuming the orbital distribution in LEO is similar to that of Starlink and OneWeb.
+A future with 400,000 LEO satellites rather than 40,000 --- the stated goal of various companies intending to launch very large constellations given the present filings\footnote{\url{https://planet4589.org/space/stats/conlist.html}} --- could  render the trade-off of 10\% of LSST observing time in order to cut the number of visits with streaks in half worthwhile.
 Satellite streaks from Starlink and OneWeb as presently designed are not expected to saturate the LSST Camera's CCD detectors.
@@
 However, it is possible for larger satellites in higher orbits, where they appear more in focus --- like OneWeb at 1200 km --- to leave faint scattered light artifacts that may be visible to the LSST Camera.
-This possibility notwithstanding, even with all three simulated constellations in orbit, the percentage of pixels lost is relatively small.
-For comparison, the Rubin focal plane has a fill factor of 88.0\%\footnote{\url{https://smtn-004.lsst.io}}, meaning 12\% of potential observations are lost to chip gaps, and losses due to weather can vary by up to 8\%.
-We also expect diffraction spikes from bright stars to render $\sim 1-2$\% of pixels in each exposure unusable.
+% I'll let it go, but if I was a hostile referee I would demand a citation for this claim.
+I've never seen an image with an unilluminated satellite in it.
+ In addition, satellites from other operators may be significantly brighter than present-day Starlink and OneWeb satellites, and may saturate the LSST Camera’s detectors or cause overwhelming levels of non-linear crosstalk.
 
-While the satellite avoidance algorithm we have presented is mostly unnecessary for the three satellite constellations considered, there is still the likelihood for a sharp increases in satellite population in the next $5–10$ years, overlapping the LSST operations period ($2024–2035$).
-With a dramatic increase in satellite population, the ability to avoid satellites might become more relevant.
-In addition, satellites from other operators may be significantly brighter than present-day Starlink and OneWeb satellites, and may saturate the LSST Camera’s detectors or cause overwhelming levels of non-linear crosstalk.
-In this case, avoidance may be required as the number of pixels lost per streak would be much higher.
-
-Related to this, one future work direction involves adding brightness weighting to the satellite avoidance algorithm.
+Therefore, one future work direction involves adding brightness weighting to the satellite avoidance algorithm.
 The idea is to only avoid satellites brighter than a certain brightness threshold.
-This could potentially reduce the region of avoidance, therefore reducing the loss in co-added depth --- it could reduce the trade off between pixel loss and loss in survey depth, making this preliminary work directly relevant for a successful LSST.
+This could potentially reduce the region of avoidance, therefore reducing the loss in observing time and co-added depth.
 It may be possible to compute optimal starting locations for a series of observations based on satellite forecasts to further optimize satellite avoidance.
@@
 
-\bibliography{ms}{}
-\bibliographystyle{aasjournal}
-
 \begin{acknowledgments}
@@
 
+\bibliography{ms}{}
+\bibliographystyle{aasjournal}
+
 \end{document}

### 2022-11-23 ms.tex
revised

@@
 Overall, sacrificing 10\% of LSST observing time to avoid satellites reduces the fraction of LSST visits with streaks by a factor of two.
-Whether such mitigation will be required depends on the overall impact of streaks on science which is not well quantified yet due to a lack of adequate information about satellite brightness distribution and the impact of glints and low surface brightness residuals on alert purity
-and systematic errors in cosmological parameter estimation.
+Whether such a mitigation will be required depends on the overall impact of streaks on science, which is not yet well quantified.
+This is due to a lack of adequate information about satellite brightness distributions as well as the impact of glints and low surface brightness residuals on alert purity and systematic errors in cosmological parameter estimation.
 A significant increase in the number of satellites or their brightness during Rubin Operations may make implementing this satellite avoidance strategy worthwhile.
@@
 
-Rubin Observatory's LSST is a ten-year astronomical imaging survey that will begin in 2024 from a new telescope under construction in Chile.
+Rubin Observatory's Legacy Survey of Space and Time (LSST) is a ten-year astronomical imaging survey that will begin in 2024 from a new telescope under construction in Chile.
 Instead of soliciting individual requests for what the telescope should observe, the LSST will uniformly survey the sky every few nights using six color filters to create a decade-long high-resolution survey of the entire southern sky, and share massive quantities of data products with the astronomy community \citep{overview}.
 To accomplish this, the LSST will employ a scheduling algorithm that uses a modified Markov Decision Process which can generate lists of desirable observations in real time \citep{naghib19}.
-The Rubin scheduler balances the desire to minimize slew time, optimize signal to noise in individual images, and to maintain survey footprint uniformity.
+The LSST scheduler balances the desire to minimize slew time, optimize signal to noise in individual images, and to maintain survey footprint uniformity.
 
@@
 LEO satellites are visible from Earth because they reflect sunlight, especially during twilight.
-As the sun-illuminated satellites move across the field of view of an astronomical pointing, they leave a streak in the image.
+As the Sun-illuminated satellites move across the field of view of an astronomical pointing, they leave a streak in the image.
 While the flux from satellite streaks can in many cases be identified and removed, the resulting pixels have much lower signal-to-noise.
@@
 They concluded that attempting to naively dodge of order 48,000 LEO satellites is useless because it is operationally inefficient.
-In this paper, rather than try to avoid all satellite streaks, we incorporate satellite avoidance as a component of the Rubin scheduler's Markov Decision Process.
+In this paper, rather than try to avoid all satellite streaks, we incorporate satellite avoidance as a component of the LSST scheduler's Markov Decision Process.
 This allows us to avoid a significant fraction of satellite streaks and investigate what level of avoidance might be acceptable because it does not drastically impact the overall performance of the LSST.
@@
 The rapid increase in population of LEO satellites threatens to compromise the quality and scientific value of LSST images and also requires extra human and computer resources to effectively mask trails.
-Thus, we explore an additional option: incorporating the orbits of known commercial satellites into the Rubin scheduler so the worst of them may be avoided.
+Thus, we explore an additional option: incorporating the orbits of known commercial satellites into the LSST scheduler so the worst of them may be avoided.
 
@@
 We then build a tool that uses that data to create new scheduler constraints, and test the impact of the modified scheduler algorithm on LSST observing programs in Section \ref{results}.
-Finally, we discuss the resulting trade-off between pixel loss and reduced survey depth that results from avoiding satellites in Section \ref{discuss}, and lay out possibilities for the future as the satellite population changes during Rubin Operations.
+Finally, we discuss the resulting trade-off between number of streaks and reduced survey depth that results from avoiding satellites in Section \ref{discuss}, and lay out possibilities for the future as the satellite population changes during Rubin Operations.
 We also make available a GitHub repository with the data and software necessary to reproduce the paper's figures\footnote{\url{https://github.com/lsst-sims/satellite-dodging-ApJL}}.
@@
 
-To simulate Rubin observations, we start with the baseline observing strategy in \citet{yoachim2022b}.
-This baseline attempts to take most observations in mixed filter pairs (e.g., $r$\ followed by $i$ 33 minutes later), and completes 215,000 visits in the first year.
-Rubin uses visits of one 30s exposure in $u$, and two back-to-back 15s exposures in the $grizy$ filters.
+To simulate LSST observations, we start with the baseline observing strategy in \citet{yoachim2022b}.
+LSST observations are scheduled in visits, where a $u$ visit is one 30s exposure and visits in all other filters ($grizy$) are back-to-back 15s exposures.
+The baseline strategy attempts to take most observations in mixed filter pairs (e.g., an $r$\ visit followed by an $i$ visit 33 minutes later), and completes 215,000 visits in the first year.
 
-The baseline Rubin observing strategy uses three primary basis functions which reward (1) minimizing slewtime, (2) maximizing the depth of images (e.g., by avoiding the Moon and high airmass), and (3) maintaining a uniform survey footprint.
+The baseline LSST observing strategy uses three primary basis functions which reward (1) minimizing slewtime, (2) maximizing the depth of images (e.g., by avoiding the Moon and high airmass), and (3) maintaining a uniform survey footprint.
 To this, we add a fourth basis function which penalizes observing areas of the sky which will have high concentrations of illuminated satellites.
@@
 
-We show three example satellite avoidance maps ready for use by the Rubin scheduler in Figure \ref{fig-simulated-scheduler}.
+We show three example satellite avoidance maps ready for use by the LSST scheduler in Figure \ref{fig-simulated-scheduler}.
 It is apparent from Figure \ref{fig-simulated-scheduler} that the simulated OneWeb constellation has more negative area --- regions that should be avoided due to large numbers of illuminated satellites --- than the other two constellations.
-Although OneWeb has fewer satellites than Starlink Gen2, the OneWeb satellites orbit at a higher altitude, meaning that they will be illuminated for a longer portion of the night, and also have a larger impact close to twilight.
+Although OneWeb has fewer satellites than Starlink Gen2, the OneWeb satellites orbit at a higher altitude (1200 km compared to 340-614 km for Starlink), meaning that they will be illuminated for a longer portion of the night, and also have a larger impact close to twilight.
 This is why one of the recommendations from \citet{satcon1} is to keep LEO satellites below 600 km altitude.
@@
 To estimate the fraction of pixels affected by streaks, following \citet{hasan22} we adopt a fiducial width of 300 pixels (equivalent to 1 arcminute given the plate scale of 0.2 arcseconds per pixel).
-Assuming a length of 15 CCDs (with a CCD side of 4096 pixels), a single streak
-corresponds to 0.6\% of all the pixels in the focal plane. On the other hand, if a streak
-is so bright that entire CCDs are rendered scientifically useless, a single streak
-would wipe out 8\% of all pixels in the focal plane.
-At this time, the overall impact of streaks on science cannot be well quantified due to a lack of adequate information about satellite brightness distribution and the impact of glints and low surface brightness residuals on alert purity and systematic errors in cosmological parameter estimation.
-That is, not only that the fraction of affected pixels is an imperfect metric for 
-assessing the impact of streaks on science, but even this fraction cannot be yet
-well determined in an absolute sense because the effective streak width depends on unknown satellite brightness.
+The 3.5 degree diameter Rubin focal plane is populated with 189 4kx4k CCDs.
+Assuming a length of 15 CCDs (with a CCD side of 4096 pixels), a single streak corresponds to 0.6\% of all the pixels in the focal plane.
+On the other hand, if a streak is so bright that entire CCDs are rendered scientifically useless, a single streak would wipe out 8\% of all pixels in the focal plane.
 
-We simulate observations for only the first year of the planned ten-year LSST.
-The survey strategy does not significantly change in later years.
+We simulate observations for only the first year of the planned ten-year LSST, as the survey strategy does not significantly change in later years.
 We acknowledge this does not account for the likely satellite population increase beyond the three simulated constellations; however, our results should scale linearly to larger future constellations in similar orbital distributions.
@@
 We estimate that satellite orbital solutions correct to within about a degree in space and to within a few minutes in time would be sufficient to effectively avoid some regions of the sky with more satellites in large constellations.
+Observing current Starlink satellites, \citet{Halferty2022} find they can construct TLEs which predict positions with sub-degree spatial precision and sub-second temporal precision which is more than adequate for our proposed satellite avoidance methods.
 
@@
 One important LSST survey goal is to collect a large number of exposures of the whole southern sky so these may be co-added to reveal faint structures that are not visible in individual visits.
-Therefore, survey depth is crucial to LSST science, and the trade-off between pixel loss from satellite streaks versus survey depth reduction from fewer total visits must be evaluated.
+As a result, survey depth is crucial to LSST science, and the trade-off between pixel loss from satellite streaks versus survey depth reduction from fewer total visits must be evaluated.
 With the satellite avoidance algorithm, the scheduler is prompted to avoid regions with illuminated satellites, which sometimes results in longer slew times or less desirable pointing conditions and contributes to a loss in survey depth.
@@
 \begin{figure}[ht!]
-\plotone{plots/streak_length_v_deltam.pdf}
+\plotone{plots/streak_length_v_streakl.pdf}
 \epsscale{1.5}
@@
 Because a small potential area is targeted, our proposed satellite avoidance scheme is ineffective.
-Figure~\ref{fig:twi_neo} shows how regular survey observations and twilight NEO observations would be affected by satellite constellations using the Rubin scheduler with no satellite avoidance.
-While the majority of twilight NEO observations could include a satellite streak with a Starlink Gen2 size constellation, we find this would only result in a $\sim$0.36\% loss of science pixels.
+Figure~\ref{fig:twi_neo} shows how regular survey observations and twilight NEO observations would be affected by satellite constellations using the LSST scheduler with no satellite avoidance.
+While the majority of twilight NEO observations could include a satellite streak with a Starlink Gen2 size constellation, we find this would only result in a $\sim$0.36\% loss of science pixels (for a 1 arcminute wide mask).
 
@@
 Compared to standard LSST observations (top), twilight NEO observations (bottom) cannot easily be shifted to avoid satellites.
-The left panels show the altitude and azimuth distribution of observations on the sky, and the right panels show how many streaks would result from the three simulated satellite constellations as a function of how high above the horizon the telescope is pointing (observation altitude).
+The left panels show the altitude and azimuth distribution of observations on the sky (zenith at the center of the plots), and the right panels show how many streaks would result from the three simulated satellite constellations as a function of how high above the horizon the telescope is pointing (observation altitude).
 Note that  most twilight NEO observations would contain a satellite streak.
@@
 The trade-off essentially comes down to the relationship between total streak length reduction and survey depth reduction.
-The final decision will depend on the overall impact of streaks on science which is not well quantified yet due to a lack of adequate information about satellite brightness distribution and the impact of glints and low surface brightness residuals on alert purity and systematic errors in cosmological parameter estimation.
+The final decision will depend on the overall impact of streaks on science which is not well quantified yet due to a lack of adequate information about the satellite brightness distribution and the impact of glints and low surface brightness residuals on alert purity and systematic errors in cosmological parameter estimation.
 In other words,
@@
 
-We note that earlier publications \citep{lawrence22,tyson20}, stated that the majority of LSST images are likely to contain a satellite streak.
+We note that earlier publications \citep{lawrence22,tyson20} stated that the majority of LSST images are likely to contain a satellite streak.
 They also included some higher-altitude Starlink orbits that are no longer planned.
@@
 A future with 400,000 LEO satellites rather than 40,000 --- the stated goal of various companies intending to launch very large constellations given the present filings\footnote{\url{https://planet4589.org/space/stats/conlist.html}} --- could render the trade-off of 10\% of LSST observing time in order to cut the number of visits with streaks in half worthwhile.
-Satellite streaks from Starlink and OneWeb as presently designed are not expected to saturate the LSST Camera's CCD detectors.
-Satellites typically only leave streaks in images when they are both illuminated by the Sun and visible from the observatory.
-LEO satellites spend most of the night in Earth's shadow.
-However, it is possible for larger satellites in higher orbits, where they appear more in focus --- like OneWeb at 1200 km --- to leave faint scattered light artifacts that may be visible to the LSST Camera.
-% I'll let it go, but if I was a hostile referee I would demand a citation for this claim.
-I've never seen an image with an unilluminated satellite in it.
- In addition, satellites from other operators may be significantly brighter than present-day Starlink and OneWeb satellites, and may saturate the LSST Camera’s detectors or cause overwhelming levels of non-linear crosstalk.
+\citet{tyson20} find that satellites with AB magnitudes of $g \sim 3.2$ to $y \sim 1.5$ would saturate LSST images. 
+Satellite streaks from Starlink and OneWeb as presently designed are not expected to saturate the LSST Camera's CCD detectors as they have magnitudes of $\sim5$ in their final orbits \citep{Halferty2022}.
+While a star of $g=10$ would be very saturated in a LSST images, LEO satellites are moving fast enough that their effective exposure time is much lower than astronomical targets.
+Satellites typically only leave streaks in images when they are both illuminated by the Sun and visible from the observatory, and LEO satellites spend most of the night in Earth's shadow.
+Satellites from other operators may be significantly brighter than present-day Starlink and OneWeb satellites, and may saturate the LSST Camera’s detectors or cause overwhelming levels of non-linear crosstalk.
+In particular, the Blue Walker 3 satellite is the first of a proposed 100 satellite constellation which have a $V$ magnitude between 1-0.
+Such a bright satellite would saturate LSST images, potentially causing much higher pixel losses than satellites which have been launched to date.
 
@@
 \begin{acknowledgments}
-JAH acknowledges support from the Computing Research Association-Widening Participation (CRA-WP) Distributed Research Experiences for Undergraduates (DREU) program.
-JAH and MLR are grateful for LSST Corporation travel support for JAH to attend the 2022 Rubin Project and Community Workshop and present a poster.
+We thank the anonymous referee who provided helpful prompt feedback that improved the clarity of the paper.
+JH acknowledges support from the Computing Research Association-Widening Participation (CRA-WP) Distributed Research Experiences for Undergraduates (DREU) program.
+JH and MLR are grateful for LSST Corporation travel support for JH to attend the 2022 Rubin Project and Community Workshop and present a poster.
 MLR acknowledges support from Bob Blum, Leanne Guy, and all of Rubin Operations to spend a fraction of her time on satellite mitigation work; this study would not have been possible without formal recognition that the proliferation of bright commercial LEO satellites poses a threat to LSST science.

### 2022-11-28 for_ph/ms.tex
ph version

--- 2022-11-23 ms.tex
@@
 \epsscale{0.37}
-\plotone{plots/starlink_tles_v1_0.00_basisfunc.pdf}
-\plotone{plots/starlink_tles_v2_0.00_basisfunc.pdf}
-\plotone{plots/oneweb_tles_0.00_basisfunc.pdf}
+\plotone{plots/starlink_tles_v1_0_00_basisfunc.pdf}
+\plotone{plots/starlink_tles_v2_0_00_basisfunc.pdf}
+\plotone{plots/oneweb_tles_0_00_basisfunc.pdf}
 \epsscale{1}
@@
 
-\bibliography{ms}{}
 \bibliographystyle{aasjournal}
+\begin{thebibliography}{}
+\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
+\providecommand{\url}[1]{\href{#1}{#1}}
+\providecommand{\dodoi}[1]{doi:~\href{http://doi.org/#1}{\nolinkurl{#1}}}
+\providecommand{\doeprint}[1]{\href{http://ascl.net/#1}{\nolinkurl{http://ascl.net/#1}}}
+\providecommand{\doarXiv}[1]{\href{https://arxiv.org/abs/#1}{\nolinkurl{https://arxiv.org/abs/#1}}}
+
+\bibitem[{{Astropy Collaboration} {et~al.}(2013){Astropy Collaboration},
+  {Robitaille}, {Tollerud}, {Greenfield}, {Droettboom}, {Bray}, {Aldcroft},
+  {Davis}, {Ginsburg}, {Price-Whelan}, {Kerzendorf}, {Conley}, {Crighton},
+  {Barbary}, {Muna}, {Ferguson}, {Grollier}, {Parikh}, {Nair}, {Unther},
+  {Deil}, {Woillez}, {Conseil}, {Kramer}, {Turner}, {Singer}, {Fox}, {Weaver},
+  {Zabalza}, {Edwards}, {Azalee Bostroem}, {Burke}, {Casey}, {Crawford},
+  {Dencheva}, {Ely}, {Jenness}, {Labrie}, {Lim}, {Pierfederici}, {Pontzen},
+  {Ptak}, {Refsdal}, {Servillat}, \& {Streicher}}]{astropy2013}
+{Astropy Collaboration}, {Robitaille}, T.~P., {Tollerud}, E.~J., {et~al.} 2013,
+  \aap, 558, A33, \dodoi{10.1051/0004-6361/201322068}
+
+\bibitem[{{Astropy Collaboration} {et~al.}(2018){Astropy Collaboration},
+  {Price-Whelan}, {Sip{\H{o}}cz}, {G{\"u}nther}, {Lim}, {Crawford}, {Conseil},
+  {Shupe}, {Craig}, {Dencheva}, {Ginsburg}, {VanderPlas}, {Bradley},
+  {P{\'e}rez-Su{\'a}rez}, {de Val-Borro}, {Aldcroft}, {Cruz}, {Robitaille},
+  {Tollerud}, {Ardelean}, {Babej}, {Bach}, {Bachetti}, {Bakanov}, {Bamford},
+  {Barentsen}, {Barmby}, {Baumbach}, {Berry}, {Biscani}, {Boquien}, {Bostroem},
+  {Bouma}, {Brammer}, {Bray}, {Breytenbach}, {Buddelmeijer}, {Burke},
+  {Calderone}, {Cano Rodr{\'\i}guez}, {Cara}, {Cardoso}, {Cheedella}, {Copin},
+  {Corrales}, {Crichton}, {D'Avella}, {Deil}, {Depagne}, {Dietrich}, {Donath},
+  {Droettboom}, {Earl}, {Erben}, {Fabbro}, {Ferreira}, {Finethy}, {Fox},
+  {Garrison}, {Gibbons}, {Goldstein}, {Gommers}, {Greco}, {Greenfield},
+  {Groener}, {Grollier}, {Hagen}, {Hirst}, {Homeier}, {Horton}, {Hosseinzadeh},
+  {Hu}, {Hunkeler}, {Ivezi{\'c}}, {Jain}, {Jenness}, {Kanarek}, {Kendrew},
+  {Kern}, {Kerzendorf}, {Khvalko}, {King}, {Kirkby}, {Kulkarni}, {Kumar},
+  {Lee}, {Lenz}, {Littlefair}, {Ma}, {Macleod}, {Mastropietro}, {McCully},
+  {Montagnac}, {Morris}, {Mueller}, {Mumford}, {Muna}, {Murphy}, {Nelson},
+  {Nguyen}, {Ninan}, {N{\"o}the}, {Ogaz}, {Oh}, {Parejko}, {Parley}, {Pascual},
+  {Patil}, {Patil}, {Plunkett}, {Prochaska}, {Rastogi}, {Reddy Janga},
+  {Sabater}, {Sakurikar}, {Seifert}, {Sherbert}, {Sherwood-Taylor}, {Shih},
+  {Sick}, {Silbiger}, {Singanamalla}, {Singer}, {Sladen}, {Sooley},
+  {Sornarajah}, {Streicher}, {Teuben}, {Thomas}, {Tremblay}, {Turner},
+  {Terr{\'o}n}, {van Kerkwijk}, {de la Vega}, {Watkins}, {Weaver}, {Whitmore},
+  {Woillez}, {Zabalza}, \& {Astropy Contributors}}]{astropy2018}
+{Astropy Collaboration}, {Price-Whelan}, A.~M., {Sip{\H{o}}cz}, B.~M., {et~al.}
+  2018, \aj, 156, 123, \dodoi{10.3847/1538-3881/aabc4f}
+
+\bibitem[{{Astropy Collaboration} {et~al.}(2022){Astropy Collaboration},
+  {Price-Whelan}, {Lim}, {Earl}, {Starkman}, {Bradley}, {Shupe}, {Patil},
+  {Corrales}, {Brasseur}, {N{\"o}the}, {Donath}, {Tollerud}, {Morris},
+  {Ginsburg}, {Vaher}, {Weaver}, {Tocknell}, {Jamieson}, {van Kerkwijk},
+  {Robitaille}, {Merry}, {Bachetti}, {G{\"u}nther}, {Aldcroft},
+  {Alvarado-Montes}, {Archibald}, {B{\'o}di}, {Bapat}, {Barentsen},
+  {Baz{\'a}n}, {Biswas}, {Boquien}, {Burke}, {Cara}, {Cara}, {Conroy},
+  {Conseil}, {Craig}, {Cross}, {Cruz}, {D'Eugenio}, {Dencheva}, {Devillepoix},
+  {Dietrich}, {Eigenbrot}, {Erben}, {Ferreira}, {Foreman-Mackey}, {Fox},
+  {Freij}, {Garg}, {Geda}, {Glattly}, {Gondhalekar}, {Gordon}, {Grant},
+  {Greenfield}, {Groener}, {Guest}, {Gurovich}, {Handberg}, {Hart},
+  {Hatfield-Dodds}, {Homeier}, {Hosseinzadeh}, {Jenness}, {Jones}, {Joseph},
+  {Kalmbach}, {Karamehmetoglu}, {Ka{\l}uszy{\'n}ski}, {Kelley}, {Kern},
+  {Kerzendorf}, {Koch}, {Kulumani}, {Lee}, {Ly}, {Ma}, {MacBride}, {Maljaars},
+  {Muna}, {Murphy}, {Norman}, {O'Steen}, {Oman}, {Pacifici}, {Pascual},
+  {Pascual-Granado}, {Patil}, {Perren}, {Pickering}, {Rastogi}, {Roulston},
+  {Ryan}, {Rykoff}, {Sabater}, {Sakurikar}, {Salgado}, {Sanghi}, {Saunders},
+  {Savchenko}, {Schwardt}, {Seifert-Eckert}, {Shih}, {Jain}, {Shukla}, {Sick},
+  {Simpson}, {Singanamalla}, {Singer}, {Singhal}, {Sinha}, {Sip{\H{o}}cz},
+  {Spitler}, {Stansby}, {Streicher}, {{\v{S}}umak}, {Swinbank}, {Taranu},
+  {Tewary}, {Tremblay}, {Val-Borro}, {Van Kooten}, {Vasovi{\'c}}, {Verma}, {de
+  Miranda Cardoso}, {Williams}, {Wilson}, {Winkel}, {Wood-Vasey}, {Xue},
+  {Yoachim}, {Zhang}, {Zonca}, \& {Astropy Project Contributors}}]{astropy2022}
+{Astropy Collaboration}, {Price-Whelan}, A.~M., {Lim}, P.~L., {et~al.} 2022,
+  \apj, 935, 167, \dodoi{10.3847/1538-4357/ac7c74}
+
+\bibitem[{Gillies {et~al.}(2007--)}]{shapely2007}
+Gillies, S., {et~al.} 2007--, Shapely: manipulation and analysis of geometric
+  objects.
+\newblock \url{https://github.com/Toblerity/Shapely}
+
+\bibitem[{{G{\'o}rski} {et~al.}(2005){G{\'o}rski}, {Hivon}, {Banday},
+  {Wandelt}, {Hansen}, {Reinecke}, \& {Bartelmann}}]{healpix2005}
+{G{\'o}rski}, K.~M., {Hivon}, E., {Banday}, A.~J., {et~al.} 2005, \apj, 622,
+  759, \dodoi{10.1086/427976}
+
+\bibitem[{{Halferty} {et~al.}(2022){Halferty}, {Reddy}, {Campbell}, {Battle},
+  \& {Furfaro}}]{Halferty2022}
+{Halferty}, G., {Reddy}, V., {Campbell}, T., {Battle}, A., \& {Furfaro}, R.
+  2022, \mnras, 516, 1502, \dodoi{10.1093/mnras/stac2080}
+
+\bibitem[{{Hall} {et~al.}(2021){Hall}, {Walker}, {Rawls}, {McDowell}, {Seaman},
+  {Venkatesan}, {Lowenthal}, {Green}, {Krafton}, \& {Parriott}}]{satcon2}
+{Hall}, J., {Walker}, C., {Rawls}, M., {et~al.} 2021, in Bulletin of the
+  American Astronomical Society, 2.205, \dodoi{10.3847/25c2cfeb.4554c01f}
+
+\bibitem[{Harris {et~al.}(2020)Harris, Millman, van~der Walt, Gommers,
+  Virtanen, Cournapeau, Wieser, Taylor, Berg, Smith, Kern, Picus, Hoyer, van
+  Kerkwijk, Brett, Haldane, del R{\'{i}}o, Wiebe, Peterson,
+  G{\'{e}}rard-Marchant, Sheppard, Reddy, Weckesser, Abbasi, Gohlke, \&
+  Oliphant}]{harris2020array}
+Harris, C.~R., Millman, K.~J., van~der Walt, S.~J., {et~al.} 2020, Nature, 585,
+  357, \dodoi{10.1038/s41586-020-2649-2}
+
+\bibitem[{{Hasan} {et~al.}(2022){Hasan}, {Tyson}, {Saunders}, \&
+  {Xin}}]{hasan22}
+{Hasan}, I., {Tyson}, J.~A., {Saunders}, C., \& {Xin}, B. 2022, Astronomy and
+  Computing, 39, 100584, \dodoi{10.1016/j.ascom.2022.100584}
+
+\bibitem[{Hunter(2007)}]{Hunter:2007}
+Hunter, J.~D. 2007, Computing in Science \& Engineering, 9, 90,
+  \dodoi{10.1109/MCSE.2007.55}
+
+\bibitem[{{Ivezi{\'c}} {et~al.}(2019){Ivezi{\'c}}, {Kahn}, {Tyson}, {Abel},
+  {Acosta}, {Allsman}, {Alonso}, {AlSayyad}, {Anderson}, {Andrew}, \&
+  et~al.}]{overview}
+{Ivezi{\'c}}, {\v Z}., {Kahn}, S.~M., {Tyson}, J.~A., {et~al.} 2019, \apj, 873,
+  111, \dodoi{10.3847/1538-4357/ab042c}
+
+\bibitem[{{Lawrence} {et~al.}(2022){Lawrence}, {Rawls}, {Jah}, {Boley}, {Di
+  Vruno}, {Garrington}, {Kramer}, {Lawler}, {Lowenthal}, {McDowell}, \&
+  {McCaughrean}}]{lawrence22}
+{Lawrence}, A., {Rawls}, M.~L., {Jah}, M., {et~al.} 2022, Nature Astronomy, 6,
+  428, \dodoi{10.1038/s41550-022-01655-6}
+
+\bibitem[{{Naghib} {et~al.}(2019){Naghib}, {Yoachim}, {Vanderbei}, {Connolly},
+  \& {Jones}}]{naghib19}
+{Naghib}, E., {Yoachim}, P., {Vanderbei}, R.~J., {Connolly}, A.~J., \& {Jones},
+  R.~L. 2019, \aj, 157, 151, \dodoi{10.3847/1538-3881/aafece}
+
+\bibitem[{{Rhodes}(2019)}]{Rhodes2019}
+{Rhodes}, B. 2019, {Skyfield: High precision research-grade positions for
+  planets and Earth satellites generator}, Astrophysics Source Code Library,
+  record ascl:1907.024.
+\newblock \doeprint{1907.024}
+
+\bibitem[{{Tyson} {et~al.}(2020){Tyson}, {Ivezi{\'c}}, {Bradshaw}, {Rawls},
+  {Xin}, {Yoachim}, {Parejko}, {Greene}, {Sholl}, {Abbott}, \&
+  {Polin}}]{tyson20}
+{Tyson}, J.~A., {Ivezi{\'c}}, {\v{Z}}., {Bradshaw}, A., {et~al.} 2020, \aj,
+  160, 226, \dodoi{10.3847/1538-3881/abba3e}
+
+\bibitem[{Virtanen {et~al.}(2020)Virtanen, Gommers, Oliphant, Haberland, Reddy,
+  Cournapeau, Burovski, Peterson, Weckesser, Bright, {van der Walt}, Brett,
+  Wilson, Millman, Mayorov, Nelson, Jones, Kern, Larson, Carey, Polat, Feng,
+  Moore, {VanderPlas}, Laxalde, Perktold, Cimrman, Henriksen, Quintero, Harris,
+  Archibald, Ribeiro, Pedregosa, {van Mulbregt}, \& {SciPy 1.0
+  Contributors}}]{2020SciPy-NMeth}
+Virtanen, P., Gommers, R., Oliphant, T.~E., {et~al.} 2020, Nature Methods, 17,
+  261, \dodoi{10.1038/s41592-019-0686-2}
+
+\bibitem[{{Walker} {et~al.}(2020{\natexlab{a}}){Walker}, {Hall}, {Allen},
+  {Green}, {Seitzer}, {Tyson}, {Bauer}, {Krafton}, {Lowenthal}, {Parriott},
+  {Puxley}, {Abbott}, {Bakos}, {Barentine}, {Bassa}, {Blakeslee}, {Bradshaw},
+  {Cooke}, {Devost}, {Galad{\'\i}-Enr{\'\i}quez}, {Haase}, {Hainaut},
+  {Heathcote}, {Jah}, {Krantz}, {Kucharski}, {McDowell}, {Mr{\'o}z}, {Otarola},
+  {Pearce}, {Rawls}, {Saunders}, {Seaman}, {Siminski}, {Snyder},
+  {Storrie-Lombardi}, {Tregloan-Reed}, {Wainscoat}, {Williams}, \&
+  {Yoachim}}]{satcon1}
+{Walker}, C., {Hall}, J., {Allen}, L., {et~al.} 2020{\natexlab{a}}, in Bulletin
+  of the American Astronomical Society, Vol.~52, 0206,
+  \dodoi{10.3847/25c2cfeb.346793b8}
+
+\bibitem[{{Walker} {et~al.}(2020{\natexlab{b}}){Walker}, {Di Pippo},
+  {Aub{\'e}}, {Barentine}, {Benkhaldoun}, {Benvenuti}, {Bouroussis}, {Green},
+  {Hearnshaw}, {Mackenzie}, {Liszt}, {Lowenthal}, {Mu{\~n}oz-Tu{\~n}{\'o}n},
+  {Nield}, {Ricard}, {Rodriguez Espinosa}, {Sanhueza}, {Varela P{\'e}rez},
+  {Williams}, {Lau}, {Mu{\~n}oz Tu{\~n}{\'o}n}, {Rodr{\'\i}guez Espinosa},
+  {Mu{\~n}oz-Tu{\~n}on}, {Armas}, {Benn}, {Blattner}, {Diaz-Castro}, {Donners},
+  {Douglas}, {Ga{\v{s}}parovsk{\'y}}, {Hall}, {Koll{\'a}th}, {Novak},
+  {Ortolani}, {Rodr{\'\i}guez Espinosa}, {Lowenthal}, {Bara}, {J{\"a}gerbrand},
+  {Jechow}, {Longcore}, {Motta}, {Schlangen}, {Schroer}, {Lucatello},
+  {Galad{\'\i}-Enr{\'\i}quez}, {Micheva}, {Otarola}, {Allen}, {Bannister},
+  {Barstow}, {Bassa}, {Davies}, {Devost}, {Hainaut}, {Jah}, {Krafton},
+  {Krantz}, {Kucharski}, {Maclay}, {Massey}, {Mudd}, {McDowell}, {Parriott},
+  {Pearce}, {Rawls}, {Rotola}, {Seaman}, {Seitzer}, {Siminski},
+  {Storrie-Lombardi}, {Street}, {Tregloan-Reed}, {Tyson}, {Unda-Sanzana},
+  {Wainscoat}, \& {Zamora}}]{dqs1}
+{Walker}, C., {Di Pippo}, S., {Aub{\'e}}, M., {et~al.} 2020{\natexlab{b}},
+  {Dark \& Quiet Skies I (2020)}, Dark \& Quiet Skies I (2020), Report of the
+  conference held 5-9 October, 2020., \dodoi{10.5281/zenodo.5898785}
+
+\bibitem[{{Walker} {et~al.}(2021){Walker}, {Di Pippo}, {Aub{\'e}}, {Barentine},
+  {Benkhaldoun}, {Benvenuti}, {Bouroussis}, {Di Vruno}, {Green}, {Hearnshaw},
+  {Aoraki}, {Liszt}, {Lowenthal}, {Mu{\~n}oz-Tu{\~n}{\'o}n}, {Nield}, {Ohishi},
+  {Ricard}, {Rodriguez Espinosa}, {Rotola}, {Sanhueza}, {Var{\'e}la P{\'e}rez},
+  {Williams}, {Cameron}, {H{\"a}nel}, {Hartley}, {Krop-Benesch}, {Lien},
+  {Motta}, {Ribas}, {Rol}, {Yakushina}, {Zielinska-Dabkowska}, {Zhao},
+  {Aganaba}, {Boley}, {Cirkovic}, {Falle}, {Freeland}, {Hallet}, {Hanlon},
+  {Jakhu}, {Jamschon MacGarry}, {Johnson}, {Kazlouskaya}, {Martin},
+  {Masson-Zwaan}, {Mishra}, {Moro-Aguilar}, {Mudd}, {Simon-Butler}, {Van Eijk},
+  {Venn}, {Wager}, {Alis}, {Berry}, {Bharadwaj}, {Davis}, {Lagadec}, {Le
+  Breuilly}, {Lucatello}, {Massey}, {Parriott}, {Valdivia}, {Weeden}, {Hofer},
+  {Andrade}, {Askari}, {Baruch}, {Bohnlein}, {Dunn}, {Grunsfeld}, {Jansson},
+  {Knox}, {Raval}, {Sedwick}, {Smith}, {Vanotti}, {Bassa}, {Cui}, {Galadi},
+  {Hainaut}, {Hasan}, {K{\'e}b{\'e}}, {Lawler}, {McDowell}, {Otarola}, {Rawls},
+  {Seitzer}, {Street}, {Pree}, {Gergely}, {Hellbourg}, {Indermuehle},
+  {Lindqvist}, {Tzioumis}, {Vanderley}, {Winkel}, \& {Zhang}}]{dqs2}
+---. 2021, {Dark \& Quiet Skies II (2021)}, Dark \& Quiet Skies II (2021),
+  Report of the conference held 3-7 October, 2021.,
+  \dodoi{10.5281/zenodo.5874725}
+
+\bibitem[{Yoachim {et~al.}(2022{\natexlab{a}})Yoachim, Jones, Eric H.~Neilsen,
+  \& Ribeiro}]{yoachim2022b}
+Yoachim, P., Jones, R.~L., Eric H.~Neilsen, J., \& Ribeiro, T.
+  2022{\natexlab{a}}, {lsst-sims/sims\_featureScheduler\_runs2.2: Initial
+  release}, v1.0.0,  Zenodo, \dodoi{10.5281/zenodo.7150784}
+
+\bibitem[{Yoachim {et~al.}(2022{\natexlab{b}})Yoachim, Jones, Eric H.~Neilsen,
+  Tiago, Carlin, pgris, Prisinzano, Bellm, Sick, lmptc, LI, nsabrams, Becker,
+  Bricman, Bregeon, Kelley, \& Andreoni}]{yoachim2022}
+Yoachim, P., Jones, L., Eric H.~Neilsen, J., {et~al.} 2022{\natexlab{b}},
+  lsst/rubin\_sim: 0.12.1, 0.12.1,  Zenodo, \dodoi{10.5281/zenodo.7087823}
+
+\bibitem[{Zonca {et~al.}(2019)Zonca, Singer, Lenz, Reinecke, Rosset, Hivon, \&
+  Gorski}]{Zonca2019}
+Zonca, A., Singer, L., Lenz, D., {et~al.} 2019, Journal of Open Source
+  Software, 4, 1298, \dodoi{10.21105/joss.01298}
+
+\end{thebibliography}
 

### 2022-11-28 make_ph_zip.sh
a ph version with tar

@@ -1,0 +1,1 @@
+tar -cvfh ms.tar for_ph/* 

### 2022-11-28 make_ph_zip.sh
ugh

@@ -1,1 +1,2 @@
-tar -cvfh ms.tar for_ph/* 
+tar -chvf ms.tar aasjournal.bst aastex631.cls ms.tex plots
+
